---
title: "LLM Project: Executive Summary"
format: 
  html:
    page-layout: full
---

```{r setup, include=FALSE}
library(llm)
library(dplyr)
library(gt)
library(scales)
library(here)

# Load data for summary
daily_data <- load_cached_ccusage("daily", project_filter = NULL)
```

Welcome to the **LLM Project Dashboard**. This page provides a high-level overview of system status and resource consumption.

## Current Status

::: {.grid}

::: {.g-col-4}
### 💰 Today's Spend
```{r spend-today, echo=FALSE}
if (!is.null(daily_data) && nrow(daily_data) > 0) {
  today_cost <- daily_data |> 
    filter(date == max(date)) |> 
    summarise(cost = sum(totalCost)) |> 
    pull(cost)
  cat("#### ", dollar(today_cost))
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-4}
### 📊 Weekly Trend
```{r weekly-spend, echo=FALSE}
if (!is.null(daily_data) && nrow(daily_data) > 0) {
  weekly_cost <- daily_data |> 
    filter(as.Date(date) > Sys.Date() - 7) |> 
    summarise(cost = sum(totalCost)) |> 
    pull(cost)
  cat("#### ", dollar(weekly_cost))
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-4}
### 🛠️ Active Issues
```{r issues, echo=FALSE}
# This matches the stats from the Structure page
cat("#### 4 Open Issues")
```
:::

:::

## Quick Links

- [**LLM Usage & Costs**](vignettes/telemetry-usage.qmd): Detailed breakdown of token types and model costs.
- [**Session Efficiency**](vignettes/telemetry-efficiency.qmd): Analysis of Claude Max5 blocks and cost intensity.
- [**CI & Git Stats**](vignettes/telemetry-ci.qmd): GitHub Actions performance and commit history.
- [**Project Structure**](vignettes/telemetry-structure.qmd): File organization and repository health.

---

*Last updated: `r Sys.time()`*