---
title: "Session Efficiency"
date: "`r Sys.Date()`"
format:
  html:
    toc: false
    page-layout: full
    code-fold: true
---

{{< include _setup.qmd >}}
{{< include ../../_includes/pages-nav.qmd >}}

```{r duration-sparkline, include=FALSE}
# Pilot: derive an inline datatype sparkline from the same vig_session_metrics
# target that feeds the Duration Trends / Max5 tabs below (JohnGavin/llm#841).
# NULL/empty-safe so the pkgdown-fallback and live-store paths both degrade
# gracefully when no blocks data is available yet.
vig_metrics_for_spark <- safe_tar_read("vig_session_metrics")
dt_duration_spark <- if (!is.null(vig_metrics_for_spark) && nrow(vig_metrics_for_spark) > 0) {
  ordered <- dplyr::arrange(vig_metrics_for_spark, start)
  dt_spec(utils::tail(ordered$duration_mins, 12), "l")
} else {
  "(no data available)"
}
```

Analysis of session durations and cost efficiency, including Claude Max5 blocks. Recent session durations trended `r dt_duration_spark` over the last dozen sessions.

::: {.panel-tabset}

## Duration Trends

```{r duration-trend, fig.cap="Average Claude Code session duration over time, coloured by recency period (last week vs older), with a LOESS-smoothed trend line overlaid. The y-axis is average duration in minutes per day. Use this to see whether sessions have been getting longer or shorter recently compared to the longer-run baseline.", fig.alt="Scatter and trend line of average session duration in minutes over time, coloured by recency period."}
safe_tar_read("vig_duration_trend_plot", wrap_df = TRUE)
```

## Cost Efficiency

```{r cost-eff, fig.cap="Cost per minute of Claude Code usage over time, coloured by recency period, with a LOESS-smoothed trend line overlaid. The y-axis is cost efficiency in dollars per minute, rounded to at most three significant figures to avoid implying false precision. A rising trend means sessions are becoming more expensive per minute of use, not just more numerous.", fig.alt="Scatter and trend line of cost per minute in dollars over time, coloured by recency period."}
safe_tar_read("vig_cost_efficiency_plot", wrap_df = TRUE)
```

## Cost vs Duration

```{r cost-duration, fig.cap="Scatter plot of cost intensity in dollars per minute against session duration in minutes, coloured by recency period with a LOESS trend line overlaid. Each point is one Claude Code session block. If the trend line slopes downward, longer sessions tend to be more cost efficient per minute than shorter ones.", fig.alt="Scatter plot of cost per minute against session duration in minutes, coloured by recency period."}
safe_tar_read("vig_cost_duration_plot", wrap_df = TRUE)
```

## Model Breakdown

```{r model-breakdown, fig.height=8, fig.cap="Cost per minute over time, faceted by Claude model variant with independent y-axis scales and a LOESS trend line in each panel. This isolates whether cost efficiency differs by model rather than being dominated by whichever model is used most often. Panels with a rising trend indicate that model's sessions have become more expensive per minute over the tracked period.", fig.alt="Small-multiples scatter plot of cost per minute over time, one panel per Claude model variant."}
safe_tar_read("vig_model_session_plot", wrap_df = TRUE)
```

## Max5 Block History {#max5-block-history}

Recent history of usage blocks. "Max5" refers to the **Claude 5-hour usage limit** window^[The "Max5" metric tracks token consumption within the rolling 5-hour window enforced by Anthropic for Claude models.].

<div title="Max5 refers to Claude's 5-hour usage window where token limits are strictly enforced.">
*(Hover for info: What is Max5?)*
</div>

```{r max5-table}
safe_tar_read("vig_max5_table", wrap_df = TRUE)
```

## Cost by Project (CodexBar estimate)

CodexBar has no session-grain cost export at all (confirmed directly against
the installed CLI while investigating [#877](https://github.com/JohnGavin/llm/issues/877)) --
its `cost` command output carries at most a date and a per-model breakdown,
never a session or project identifier. The plot below is therefore a
**day/project-grain estimate**, not a per-session measurement: it apportions
CodexBar's daily total cost across projects using session-duration weighting
against `unified.duckdb.sessions`, then sums to a project total over
whatever date window the export currently covers. It is a different data
source from the Max5-block panels above (which remain fresh, per-block, and
unrelated to CodexBar) and is included here as a forward-looking
supplement, not a replacement.

```{r codexbar-project-cost, fig.cap="Estimated total cost per project over the currently available CodexBar export window, apportioned from CodexBar's day-level cost total via session-duration weighting. This is a day/project-grain estimate, not a per-session measurement -- CodexBar's raw CLI output has no session or project identifier.", fig.alt="Cleveland dot plot of estimated total cost in dollars per project, based on CodexBar's daily cost apportioned by session-duration share."}
safe_tar_read("vig_codexbar_project_cost_plot", wrap_df = TRUE)
```

:::
