---
title: "Session Efficiency"
date: "`r Sys.Date()`"
format:
  html:
    toc: false
    page-layout: full
    code-fold: true
---

{{< include _setup.qmd >}}
{{< include ../../_includes/pages-nav.qmd >}}

```{r duration-sparkline, include=FALSE}
# Pilot: derive an inline datatype sparkline from the same vig_session_metrics
# target that feeds the Duration Trends / Max5 tabs below (JohnGavin/llm#841).
# NULL/empty-safe so the pkgdown-fallback and live-store paths both degrade
# gracefully when no blocks data is available yet.
vig_metrics_for_spark <- safe_tar_read("vig_session_metrics")
dt_duration_spark <- if (!is.null(vig_metrics_for_spark) && nrow(vig_metrics_for_spark) > 0) {
  ordered <- dplyr::arrange(vig_metrics_for_spark, start)
  dt_spec(utils::tail(ordered$duration_mins, 12), "l")
} else {
  "(no data available)"
}
```

Analysis of session durations and cost efficiency, including Claude Max5 blocks. Recent session durations trended `r dt_duration_spark` over the last dozen sessions.

::: {.panel-tabset}

## Duration Trends

```{r duration-trend, fig.cap="Average session duration over time, colored by recency period."}
safe_tar_read("vig_duration_trend_plot", wrap_df = TRUE)
```

## Cost Efficiency

```{r cost-eff, fig.cap="Cost per minute over time showing session cost efficiency trends."}
safe_tar_read("vig_cost_efficiency_plot", wrap_df = TRUE)
```

## Cost vs Duration

```{r cost-duration, fig.cap="Scatter plot of cost intensity vs session duration. Longer sessions tend to be more cost efficient."}
safe_tar_read("vig_cost_duration_plot", wrap_df = TRUE)
```

## Model Breakdown

```{r model-breakdown, fig.height=8, fig.cap="Cost efficiency faceted by Claude model variant."}
safe_tar_read("vig_model_session_plot", wrap_df = TRUE)
```

## Max5 Block History {#max5-block-history}

Recent history of usage blocks. "Max5" refers to the **Claude 5-hour usage limit** window^[The "Max5" metric tracks token consumption within the rolling 5-hour window enforced by Anthropic for Claude models.].

<div title="Max5 refers to Claude's 5-hour usage window where token limits are strictly enforced.">
*(Hover for info: What is Max5?)*
</div>

```{r max5-table}
safe_tar_read("vig_max5_table", wrap_df = TRUE)
```

:::
