---
title: "From Nix Shells to Project Outputs"
format: closeread-html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Discover target store — handles non-default store in _targets.yaml
# Must setwd to project root for tar_config_get to read _targets.yaml
tar_store <- tryCatch({
  owd <- setwd(here::here())
  on.exit(setwd(owd))
  s <- targets::tar_config_get("store")
  normalizePath(s)
}, error = function(e) NULL)

safe_tar_read <- function(name) {
  if (is.null(tar_store)) return(NULL)
  tryCatch(targets::tar_read_raw(name, store = tar_store), error = function(e) NULL)
}

emit_mermaid <- function(target_name, fallback_msg = "Diagram unavailable") {
  diagram <- safe_tar_read(target_name)
  if (!is.null(diagram)) {
    id <- gsub("[^a-z0-9]", "", target_name)
    cat(sprintf('<pre class="mermaid" id="%s"></pre>\n', id))
    cat(sprintf('<script type="text/plain" data-mermaid="%s">\n', id))
    cat(diagram)
    cat("\n</script>\n")
  } else {
    cat(paste0("*", fallback_msg, "*\n"))
  }
}

counts <- safe_tar_read("vig_cr_infra_counts")
if (is.null(counts)) {
  counts <- list(
    n_rules = 58, n_skills = 68, n_agents = 12,
    n_hooks = 7, n_commands = 14, n_memory = 14
  )
}
```

<script type="module">
import mermaid from 'https://cdn.jsdelivr.net/npm/mermaid@11/dist/mermaid.esm.min.mjs';
mermaid.initialize({
  startOnLoad: false,
  securityLevel: 'loose',
  theme: 'dark',
  themeVariables: {
    background: '#000000',
    primaryColor: '#999999',
    lineColor: '#CC0000',
    primaryTextColor: '#000000'
  }
});
document.querySelectorAll('script[data-mermaid]').forEach(function(s) {
  var target = document.getElementById(s.getAttribute('data-mermaid'));
  if (target) target.textContent = s.textContent;
});
await mermaid.run({ querySelector: '.mermaid' });
</script>


## The Foundation {.cr-section}

:::{#cr-layers}

```{r layer-diagram, results='asis'}
emit_mermaid("vig_cr_layer_diagram", "Software layer diagram unavailable")
```

:::

The entire stack begins with the operating system --- macOS for development, Linux via <a href="https://orbstack.dev/" target="_blank">OrbStack</a> for CI parity. Every layer above inherits the guarantees of the layer below.

@cr-layers

<a href="https://nixos.org/" target="_blank">Nix</a> provides hermetic, reproducible package management. The <a href="https://docs.ropensci.org/rix/" target="_blank">rix</a> package translates R-flavoured declarations in `default.R` into `default.nix` derivations. Every project gets its own isolated shell with pinned package versions.

@cr-layers

Above <a href="https://nixos.org/" target="_blank">Nix</a> sit the R and Python packages specific to each project. <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> orchestrates pipelines, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> manages parallel workers, and <a href="https://shikokuchuo.net/mirai/" target="_blank">mirai</a> provides the async evaluation engine. These are the workhorses of computation.

@cr-layers

At the top, the LLM harness --- `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, and `r counts$n_memory` memory files --- guides <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> through finding, organising, and synthesising project work. The outputs flow into vignettes, APIs, and reports.

@cr-layers


## The LLM Harness {.cr-section}

:::{#cr-harness}

```{r harness-diagram, results='asis'}
emit_mermaid("vig_cr_harness_diagram", "Harness diagram unavailable")
```

:::

<a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> is the entry point --- a compressed 200-line file that <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> reads at session start. It indexes `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, `r counts$n_hooks` hooks, `r counts$n_commands` commands, and `r counts$n_memory` memory files.

@cr-harness

**Rules** enforce constraints: never use raw SQL (use <a href="https://duckdblabs.github.io/duckplyr/" target="_blank">duckplyr</a>), never hardcode values in prose, never delete files without checking size. They fire automatically based on file path globs in their YAML frontmatter.

@cr-harness

**Skills** provide deep knowledge on demand: how to set up a <a href="https://nixos.org/" target="_blank">Nix</a> environment, how to write <a href="https://testthat.r-lib.org/" target="_blank">testthat</a> tests, how to build a <a href="https://posit-dev.github.io/r-shinylive/" target="_blank">Shinylive</a> vignette. They are invoked when the task matches their domain.

@cr-harness

**Agents** are specialised subprocesses: `r-debugger` for test failures, `targets-runner` for pipelines, `critic` for read-only review. Each is pinned to a model tier --- <a href="https://docs.anthropic.com/en/docs/about-claude/models#claude-4-haiku" target="_blank">haiku</a> for typos, <a href="https://docs.anthropic.com/en/docs/about-claude/models#claude-4-sonnet" target="_blank">sonnet</a> for debugging, <a href="https://docs.anthropic.com/en/docs/about-claude/models#claude-4-opus" target="_blank">opus</a> for architecture.

@cr-harness

**Hooks** are shell scripts triggered by events: `session_init.sh` checks <a href="https://nixos.org/" target="_blank">Nix</a> and burn rate at startup, `file_protection.sh` blocks edits to `raw/` knowledge base files, `context_survival.sh` saves state before compaction.

@cr-harness

**Memory** persists across conversations: user preferences, feedback corrections, project state, and reference pointers to external systems. It is file-based, indexed by `MEMORY.md`, and updated incrementally.

@cr-harness


## irishbuoys: A Real Pipeline {.cr-section}

:::{#cr-pipeline}

```{r pipeline-diagram, results='asis'}
emit_mermaid("vig_cr_pipeline_dag", "Pipeline diagram unavailable")
```

:::

The <a href="https://github.com/JohnGavin/irishbuoys" target="_blank">irishbuoys</a> R package monitors wave conditions at Marine Institute buoy stations around Ireland. Its <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> pipeline has 19 plan files and ~294 targets spanning every layer of the software stack.

@cr-pipeline

**Data acquisition** fetches hourly observations from the <a href="https://erddap.marine.ie/" target="_blank">ERDDAP</a> API and stores them in <a href="https://duckdb.org/" target="_blank">DuckDB</a>. This is the bottom of the pipeline --- raw data entering the system through the <a href="https://nixos.org/" target="_blank">Nix</a>-managed R environment.

@cr-pipeline

**Data validation** runs 8 targets checking temporal coverage, gaps, duplicates, freshness, and sampling frequency. If any station drops below 30% coverage, the pipeline aborts. This is where <a href="https://rstudio.github.io/pointblank/" target="_blank">pointblank</a>-style validation gates earn their keep --- catching problems before they cascade.

@cr-pipeline

**Wave analysis** is the largest plan (61 targets): rolling aggregations, extreme value analysis via <a href="https://cran.r-project.org/package=extRemes" target="_blank">extRemes</a>, spatial modelling across stations. <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> workers parallelise computation across stations.

@cr-pipeline

**Outputs** fan out from analysis into multiple formats: storm alerts trigger <a href="https://pkgs.rstudio.com/blastula/" target="_blank">blastula</a> emails, <a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a> dashboards with <a href="https://plotly.com/r/" target="_blank">plotly</a> charts, a static JSON API, and <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a> vignettes. A single `tar_make()` call produces all of them.

@cr-pipeline

**QA gates** score the pipeline on a Bronze/Silver/Gold scale before any merge. The `qa_no_raw_sql` target enforces <a href="https://duckdblabs.github.io/duckplyr/" target="_blank">duckplyr</a> over raw SQL. The `qa_vignette_compliance` target checks every vignette chunk maps to a pipeline target. Nothing ships without passing.

@cr-pipeline


## FOCUS in Action {.cr-section}

:::{#cr-focus}

```{r focus-diagram, results='asis'}
emit_mermaid("vig_cr_focus_diagram", "FOCUS diagram unavailable")
```

:::

FOCUS is how <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> navigates a project. Each phase maps to specific infrastructure components.

@cr-focus

**Find** --- Claude uses Glob, Grep, and the Explore agent to locate relevant code, files, and patterns. The `ctx.yaml` cache provides pre-indexed package signatures so Claude doesn't need to re-read documentation each session.

@cr-focus

**Organize** --- <a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> provides the map. The Plan agent designs implementation strategies. Rules like `architecture-planning` enforce a brainstorming phase before any coding begins. Tasks break work into trackable steps.

@cr-focus

**Condense** --- Memory files persist key learnings across conversations. The `context_survival.sh` hook saves state before compaction. The `ctx-yaml-cache` rule ensures package context is pre-computed and cached.

@cr-focus

**Understand** --- The `r-debugger` agent applies scientific method debugging (Hypothesis-Experiment-Conclusion). The `critic` agent performs read-only adversarial review. The `quality-gates` skill scores work on a Bronze/Silver/Gold scale.

@cr-focus

**Synthesize** --- The `fixer` agent applies fixes from critic reports. The `targets-runner` executes pipelines. The `orchestrator-protocol` coordinates agents through verify-review-fix-score loops until quality gates pass.

@cr-focus


## The Pipeline Layers {.cr-section}

:::{#cr-detail}

```{r pipeline-detail-diagram, results='asis'}
emit_mermaid("vig_cr_pipeline_dag_mermaid", "Pipeline detail unavailable")
```

:::

Each plan file in the <a href="https://github.com/JohnGavin/irishbuoys" target="_blank">irishbuoys</a> pipeline touches specific software layers. This table shows what each plan does and which packages it relies on.

@cr-detail

```{r pipeline-table, results='asis'}
walkthrough <- safe_tar_read("vig_cr_pipeline_walkthrough")
if (!is.null(walkthrough)) {
  knitr::kable(walkthrough, format = "html",
    caption = paste0("irishbuoys pipeline: ", nrow(walkthrough),
                     " plan files across data, QA, analysis, and output layers."))
}
```

The pipeline starts at the **data layer**: <a href="https://erddap.marine.ie/" target="_blank">ERDDAP</a> API calls fetch raw observations into <a href="https://duckdb.org/" target="_blank">DuckDB</a>. The <a href="https://nixos.org/" target="_blank">Nix</a> shell provides both R and the system libraries needed for HTTP and database access.

@cr-detail

The **QA layer** runs three plan files with validation and quality control targets. These check temporal coverage, physical bounds, and data freshness. The `qa_gates` plan scores the entire pipeline before any merge --- nothing deploys below Bronze.

@cr-detail

The **analysis layer** is where domain logic lives: wave statistics, cross-station correlations, spatial extreme value models. <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> workers parallelise computation, and <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> caches results so only changed branches re-execute.

@cr-detail

The **output layer** fans results into multiple formats: <a href="https://pkgs.rstudio.com/blastula/" target="_blank">blastula</a> storm alert emails, <a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a> dashboards, a static JSON API, and <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a> vignettes. A single `tar_make()` call produces all outputs from the same validated data.

@cr-detail


---

```{r build-info, results='asis'}
info <- safe_tar_read("vig_cr_build_info")
if (!is.null(info)) cat(info) else cat("*Build info unavailable*")
```
