---
title: "From Nix Shells to Project Outputs"
format:
  html:
    toc: false
    page-layout: full
    code-fold: true
    embed-resources: false
    theme:
      dark: darkly
      light: cosmo
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

safe_tar_read <- function(name) {
  store <- tryCatch({
    owd <- setwd(here::here())
    on.exit(setwd(owd))
    normalizePath(targets::tar_config_get("store"))
  }, error = function(e) NULL)

  if (!is.null(store)) {
    obj <- tryCatch(targets::tar_read_raw(name, store = store), error = function(e) NULL)
    if (!is.null(obj)) return(obj)
  }

  for (d in c("../inst/extdata/vignettes", "inst/extdata/vignettes",
              file.path(here::here(), "inst/extdata/vignettes"))) {
    rds <- file.path(d, paste0(name, ".rds"))
    if (file.exists(rds)) return(readRDS(rds))
  }
  NULL
}

counts <- safe_tar_read("vig_cr_infra_counts")
if (is.null(counts)) {
  counts <- list(
    n_rules = 58, n_skills = 68, n_agents = 12,
    n_hooks = 7, n_commands = 14, n_memory = 14
  )
}
```


## Software Layers

```{mermaid}
%%| fig-cap: "Software stack from OS through [Nix](https://nixos.org/){target='_blank'} to project outputs. [rix](https://docs.ropensci.org/rix/){target='_blank'} translates `default.R` into `default.nix`. [targets](https://docs.ropensci.org/targets/){target='_blank'}, [crew](https://wlandau.github.io/crew/){target='_blank'}, and [mirai](https://shikokuchuo.net/mirai/){target='_blank'} orchestrate computation."
graph LR
  OS["OS: macOS / Linux"] --> Nix["Nix + rix"] --> Pkgs["R / Python<br/>packages"]
  Pkgs --> Orch["targets + crew<br/>+ mirai"]
  Orch --> LLM["LLM harness"]
  LLM --> Out["Outputs"]
```

::: {.panel-tabset}

### How It Works

::: {.callout-note}
## Nix + rix
<a href="https://nixos.org/" target="_blank">Nix</a> provides hermetic, reproducible package management. <a href="https://docs.ropensci.org/rix/" target="_blank">rix</a> translates R-flavoured declarations in `default.R` into `default.nix` derivations. Every project gets its own isolated shell with pinned package versions.
:::

::: {.callout-tip}
## Orchestration
Above Nix sit the project-specific R and Python packages. <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> orchestrates pipelines, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> manages parallel workers, and <a href="https://shikokuchuo.net/mirai/" target="_blank">mirai</a> provides the async evaluation engine.
:::

::: {.callout-important}
## LLM Harness
At the top, `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, and `r counts$n_memory` memory files guide <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> through finding, organising, and synthesising project work.
:::

:::


## LLM Harness

```{mermaid}
%%| fig-cap: "[AGENTS.md](https://github.com/JohnGavin/llm/blob/main/AGENTS.md){target='_blank'} is the entry point — indexing rules, skills, agents, hooks, commands, and memory files."
graph LR
  A["AGENTS.md<br/>Entry point"] --> R["Rules"]
  A --> S["Skills"]
  A --> G["Agents"]
  A --> H["Hooks"]
  A --> C["Commands"]
  A --> M["Memory"]
```

::: {.panel-tabset}

### Rules

| Rule | Enforces |
|------|----------|
| `duckdplyr-not-sql` | Use <a href="https://duckdblabs.github.io/duckplyr/" target="_blank">duckplyr</a>, never raw SQL |
| `dynamic-prose-values` | No hardcoded numbers in prose |
| `safe-deletion` | Check size/age before deleting files |
| `btw-timeouts` | All R execution via `timeout` in bash |
| `verification-before-completion` | No claims without fresh evidence |
| `look-ahead-bias-prevention` | Temporal separation in backtests |

### Skills

| Category | Count | Examples |
|----------|------:|---------|
| R Package Dev | 12 | <a href="https://testthat.r-lib.org/" target="_blank">testthat</a>, <a href="https://lifecycle.r-lib.org/" target="_blank">lifecycle</a>, <a href="https://rlang.r-lib.org/" target="_blank">rlang</a> |
| Data & Analysis | 8 | <a href="https://rstudio.github.io/pointblank/" target="_blank">pointblank</a>, EDA, modelling |
| Targets & Pipelines | 4 | Pipeline spec, vignettes, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> ops |
| Shiny & Web | 7 | <a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a>, async, <a href="https://posit-dev.github.io/r-shinylive/" target="_blank">Shinylive</a> |
| Quarto & Docs | 6 | Websites, dashboards, alt text |
| DevOps & CI | 3 | GitHub Actions, <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a> |

### Agents

| Agent | Model | Trigger |
|-------|-------|---------|
| `quick-fix` | <a href="https://docs.anthropic.com/en/docs/about-claude/models" target="_blank">haiku</a> | Typos, renames, version bumps |
| `r-debugger` | <a href="https://docs.anthropic.com/en/docs/about-claude/models" target="_blank">sonnet</a> | Test failures, R CMD check |
| `critic` | sonnet | Read-only adversarial review |
| `fixer` | sonnet | Apply fixes from critic reports |
| `targets-runner` | sonnet | Pipeline execution |
| `reviewer` | sonnet | PR code review |
| `nix-env` | sonnet | Nix shell diagnostics |

### Hooks & Memory

**Hooks** — shell scripts triggered by events:

| Hook | Event | Purpose |
|------|-------|---------|
| `session_init.sh` | Session start | Check Nix, burn rate, worktrees |
| `file_protection.sh` | Pre-edit | Block edits to `raw/` knowledge base |
| `context_survival.sh` | Pre-compact | Save state before context compression |
| `context_monitor.sh` | Post-tool | Periodic burn rate check |

**Memory** — file-based persistence across conversations:

| Type | Purpose |
|------|---------|
| `user` | Role, preferences, knowledge level |
| `feedback` | Corrections and confirmed approaches |
| `project` | Ongoing work, goals, deadlines |
| `reference` | Pointers to external systems |

:::


## irishbuoys Pipeline

```{mermaid}
%%| fig-cap: "[irishbuoys](https://github.com/JohnGavin/irishbuoys){target='_blank'} monitors wave conditions at Marine Institute buoy stations. 19 plan files, ~294 [targets](https://docs.ropensci.org/targets/){target='_blank'}. Data from [ERDDAP](https://erddap.marine.ie/){target='_blank'}, stored in [DuckDB](https://duckdb.org/){target='_blank'}. Outputs: [blastula](https://pkgs.rstudio.com/blastula/){target='_blank'} alerts, [bslib](https://rstudio.github.io/bslib/){target='_blank'} dashboards, JSON API, [pkgdown](https://pkgdown.r-lib.org/){target='_blank'} vignettes."
graph LR
  acq["ERDDAP<br/>API"] --> val["Validation<br/>8 targets"] --> qc["QC<br/>bounds"]
  qc --> wave["Wave Analysis<br/>61 targets"]
  wave --> joint["Joint<br/>Analysis"] --> sev["Spatial<br/>Extremes"]
  wave --> pred["Predictions"] --> alert["Storm<br/>Alerts"] --> email["Email<br/>blastula"]
  wave --> dash["Dashboard<br/>bslib"]
  wave --> api["Static<br/>API"]
  joint --> site["pkgdown<br/>Site"]
  qa["QA Gates<br/>Bronze/Silver/Gold"] --> site
```

::: {.panel-tabset}

### Plan Files

```{r pipeline-table}
walkthrough <- safe_tar_read("vig_cr_pipeline_walkthrough")
if (!is.null(walkthrough)) {
  knitr::kable(walkthrough,
    caption = paste0("irishbuoys pipeline: ", nrow(walkthrough),
                     " plan files across data, QA, analysis, and output layers."))
}
```

### Simplified Flow

```{mermaid}
%%| fig-cap: "Simplified view: [ERDDAP](https://erddap.marine.ie/){target='_blank'} fetch → validation → wave analysis → outputs. QA gates score Bronze/Silver/Gold before merge."
graph LR
  acq["ERDDAP"] --> val["8 validation"] --> wave["61 analysis"]
  wave --> alert["Alerts + email"]
  wave --> site["Dashboard + pkgdown"]
  qa["QA gates"] --> site
```

:::


## FOCUS Framework

```{mermaid}
%%| fig-cap: "FOCUS: how [Claude Code](https://claude.com/claude-code){target='_blank'} navigates a project. Each phase maps to specific agents, rules, and skills."
flowchart LR
  F["Find<br/>Glob, Grep, Explore"] --> O["Organize<br/>AGENTS.md, Plan"]
  O --> C["Condense<br/>Memory, ctx.yaml"]
  C --> U["Understand<br/>r-debugger, critic"]
  U --> Sy["Synthesize<br/>fixer, targets-runner"]
```

::: {.panel-tabset}

### Find

| Tool | Purpose |
|------|---------|
| `Glob` | Fast file pattern matching (`**/*.R`) |
| `Grep` | Content search with regex |
| `Explore` agent | Deep codebase exploration |
| `ctx.yaml` cache | Pre-indexed package signatures |

### Organize

| Component | Role |
|-----------|------|
| <a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> | Entry point — indexes all infrastructure |
| Plan agent | Design implementation strategies |
| `architecture-planning` rule | Mandatory brainstorming before coding |

### Condense

| Component | Role |
|-----------|------|
| Memory files | Persist learnings across conversations |
| `context_survival.sh` | Save state before compaction |
| CHANGELOG.md | Record completed work, failed approaches |

### Understand

| Component | Role |
|-----------|------|
| `r-debugger` agent | Hypothesis-Experiment-Conclusion loops |
| `critic` agent | Read-only adversarial review |
| `quality-gates` skill | Bronze/Silver/Gold scoring |

### Synthesize

| Component | Role |
|-----------|------|
| `fixer` agent | Apply fixes from critic reports |
| `targets-runner` agent | Execute pipelines |
| `orchestrator-protocol` | Verify → review → fix → score loop |

:::


## Build Info

```{r build-info}
#| results: asis
info <- safe_tar_read("vig_cr_build_info")
if (!is.null(info)) cat(info) else cat("*Run `tar_make()` locally to generate build info.*")
```
