---
title: "From Nix Shells to Project Outputs"
format:
  dashboard:
    orientation: rows
    theme:
      dark: darkly
      light: cosmo
    embed-resources: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

# Read target with RDS fallback (CI has no targets store)
safe_tar_read <- function(name) {
  store <- tryCatch({
    owd <- setwd(here::here())
    on.exit(setwd(owd))
    normalizePath(targets::tar_config_get("store"))
  }, error = function(e) NULL)

  if (!is.null(store)) {
    obj <- tryCatch(targets::tar_read_raw(name, store = store), error = function(e) NULL)
    if (!is.null(obj)) return(obj)
  }

  for (d in c("../inst/extdata/vignettes", "inst/extdata/vignettes",
              file.path(here::here(), "inst/extdata/vignettes"))) {
    rds <- file.path(d, paste0(name, ".rds"))
    if (file.exists(rds)) return(readRDS(rds))
  }
  NULL
}

counts <- safe_tar_read("vig_cr_infra_counts")
if (is.null(counts)) {
  counts <- list(
    n_rules = 58, n_skills = 68, n_agents = 12,
    n_hooks = 7, n_commands = 14, n_memory = 14
  )
}
```

<script>
// Quarto's mermaid-init.js renders on window.load, but dashboard tabs/pages
// hide content. SVGs rendered while hidden get zero dimensions.
// Fix: after initial render, find any zero-width SVGs and re-render them
// when their container becomes visible.
window.addEventListener('load', function() {
  // Wait for Quarto's mermaid rendering to finish
  setTimeout(function() {
    // Store original mermaid source before it's replaced
    // (Quarto already replaced <pre> with <svg>, but we can re-render)

    function fixMermaidSvgs() {
      document.querySelectorAll('svg.mermaid-js').forEach(function(svg) {
        var rect = svg.getBoundingClientRect();
        // If SVG has zero or very small dimensions, it was rendered while hidden
        if (rect.width < 10 || rect.height < 10) {
          // Force a reflow by temporarily making parent visible
          var parent = svg.closest('.tab-pane');
          if (parent && !parent.classList.contains('active')) {
            // Temporarily show the pane to get correct dimensions
            parent.style.display = 'block';
            parent.style.visibility = 'hidden';
            parent.style.position = 'absolute';

            // Get the viewBox and set explicit dimensions
            var vb = svg.getAttribute('viewBox');
            if (vb) {
              var parts = vb.split(' ');
              if (parts.length === 4) {
                svg.setAttribute('width', parts[2] + 'px');
                svg.setAttribute('height', parts[3] + 'px');
                svg.style.maxWidth = '100%';
              }
            }

            parent.style.display = '';
            parent.style.visibility = '';
            parent.style.position = '';
          }
        }
      });
    }

    // Run fix after Quarto mermaid init completes
    fixMermaidSvgs();

    // Also fix on tab/page switch
    document.addEventListener('shown.bs.tab', function() {
      setTimeout(fixMermaidSvgs, 100);
    });
  }, 500);
});
</script>

# Software Layers

## Row

### {.tabset}

#### Layer Diagram

```{mermaid}
graph TD
  OS["OS: macOS / Linux"]
  Nix["Nix + rix<br/>Reproducible shells"]
  Pkgs["R / Python packages<br/>Project-specific"]
  Orch["targets + crew + mirai<br/>Pipeline orchestration"]
  LLM["LLM harness<br/>Rules + Skills + Agents"]
  Out["Outputs<br/>Vignettes, APIs, Reports"]
  OS --> Nix --> Pkgs --> Orch --> LLM --> Out
```

::: {.card-footer}
Software stack from OS through <a href="https://nixos.org/" target="_blank">Nix</a> to project outputs. Each layer inherits the reproducibility guarantees of the layer below. <a href="https://docs.ropensci.org/rix/" target="_blank">rix</a> translates `default.R` into `default.nix`. <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a>, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a>, and <a href="https://shikokuchuo.net/mirai/" target="_blank">mirai</a> orchestrate computation. Source: `R/tar_plans/plan_vignette_closeread.R`.
:::

#### How It Works

::: {.callout-note}
## Nix + rix
<a href="https://nixos.org/" target="_blank">Nix</a> provides hermetic, reproducible package management. <a href="https://docs.ropensci.org/rix/" target="_blank">rix</a> translates R-flavoured declarations in `default.R` into `default.nix` derivations. Every project gets its own isolated shell with pinned package versions.
:::

::: {.callout-tip}
## Orchestration
Above Nix sit the project-specific R and Python packages. <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> orchestrates pipelines, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> manages parallel workers, and <a href="https://shikokuchuo.net/mirai/" target="_blank">mirai</a> provides the async evaluation engine.
:::

::: {.callout-important}
## LLM Harness
At the top, `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, and `r counts$n_memory` memory files guide <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> through finding, organising, and synthesising project work.
:::


# LLM Harness

## Row

### {.tabset}

#### Harness Diagram

```{mermaid}
graph LR
  A["AGENTS.md<br/>Entry point"]
  R["Rules"]
  S["Skills"]
  G["Agents"]
  H["Hooks"]
  C["Commands"]
  M["Memory"]
  A --> R
  A --> S
  A --> G
  A --> H
  A --> C
  A --> M
```

::: {.card-footer}
<a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> is the entry point — a compressed 200-line file indexing `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, `r counts$n_hooks` hooks, `r counts$n_commands` commands, and `r counts$n_memory` memory files. Source: `R/tar_plans/plan_vignette_closeread.R`.
:::

#### Rules

**Rules** enforce constraints automatically based on file path globs in their YAML frontmatter.

| Rule | Enforces |
|------|----------|
| `duckdplyr-not-sql` | Use <a href="https://duckdblabs.github.io/duckplyr/" target="_blank">duckplyr</a>, never raw SQL |
| `dynamic-prose-values` | No hardcoded numbers in prose |
| `safe-deletion` | Check size/age before deleting files |
| `btw-timeouts` | All R execution via `timeout` in bash |
| `verification-before-completion` | No claims without fresh evidence |
| `look-ahead-bias-prevention` | Temporal separation in backtests |

#### Skills

**Skills** provide deep knowledge on demand — invoked when the task matches their domain.

| Category | Count | Examples |
|----------|------:|---------|
| R Package Dev | 12 | <a href="https://testthat.r-lib.org/" target="_blank">testthat</a> patterns, <a href="https://lifecycle.r-lib.org/" target="_blank">lifecycle</a>, <a href="https://rlang.r-lib.org/" target="_blank">rlang</a> |
| Data & Analysis | 8 | <a href="https://rstudio.github.io/pointblank/" target="_blank">pointblank</a> validation, EDA, modelling |
| Targets & Pipelines | 4 | Pipeline spec, vignettes, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> ops |
| Shiny & Web | 7 | <a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a>, async, <a href="https://posit-dev.github.io/r-shinylive/" target="_blank">Shinylive</a> |
| Quarto & Docs | 6 | Websites, dashboards, alt text |
| DevOps & CI | 3 | GitHub Actions, <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a> |

#### Agents

**Agents** are specialised subprocesses pinned to model tiers.

| Agent | Model | Trigger |
|-------|-------|---------|
| `quick-fix` | <a href="https://docs.anthropic.com/en/docs/about-claude/models" target="_blank">haiku</a> | Typos, renames, version bumps |
| `r-debugger` | <a href="https://docs.anthropic.com/en/docs/about-claude/models" target="_blank">sonnet</a> | Test failures, R CMD check |
| `critic` | sonnet | Read-only adversarial review |
| `fixer` | sonnet | Apply fixes from critic reports |
| `targets-runner` | sonnet | Pipeline execution |
| `reviewer` | sonnet | PR code review |
| `nix-env` | sonnet | Nix shell diagnostics |

#### Hooks & Memory

**Hooks** — shell scripts triggered by events:

| Hook | Event | Purpose |
|------|-------|---------|
| `session_init.sh` | Session start | Check Nix, burn rate, worktrees |
| `file_protection.sh` | Pre-edit | Block edits to `raw/` knowledge base |
| `context_survival.sh` | Pre-compact | Save state before context compression |
| `context_monitor.sh` | Post-tool | Periodic burn rate check |

**Memory** — file-based persistence across conversations:

| Type | Purpose |
|------|---------|
| `user` | Role, preferences, knowledge level |
| `feedback` | Corrections and confirmed approaches |
| `project` | Ongoing work, goals, deadlines |
| `reference` | Pointers to external systems |


# irishbuoys Pipeline

## Row

### {.tabset}

#### Pipeline DAG

```{mermaid}
graph TD
  acq["Data Acquisition<br/>ERDDAP API"]
  val["Data Validation<br/>8 QA targets"]
  qc["Quality Control<br/>Physical bounds"]
  wave["Wave Analysis<br/>61 targets"]
  joint["Joint Analysis<br/>Cross-station"]
  sev["Spatial Extremes<br/>EVA models"]
  pred["Predictions<br/>Forecasts"]
  alert["Storm Alerts<br/>Threshold triggers"]
  email["Email Report<br/>blastula"]
  dash["Dashboard<br/>bslib + plotly"]
  api["Static API<br/>JSON endpoints"]
  site["pkgdown Site<br/>Vignettes"]
  qa["QA Gates<br/>Bronze/Silver/Gold"]
  acq --> val --> qc --> wave
  wave --> joint --> sev
  wave --> pred --> alert --> email
  wave --> dash
  wave --> api
  joint --> site
  qa --> site
```

::: {.card-footer}
The <a href="https://github.com/JohnGavin/irishbuoys" target="_blank">irishbuoys</a> R package monitors wave conditions at Marine Institute buoy stations. 19 plan files, ~294 <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a>. Data from <a href="https://erddap.marine.ie/" target="_blank">ERDDAP</a>, stored in <a href="https://duckdb.org/" target="_blank">DuckDB</a>. Outputs: storm alerts via <a href="https://pkgs.rstudio.com/blastula/" target="_blank">blastula</a>, <a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a> dashboards, static JSON API, <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a> vignettes.
:::

#### Plan Files

```{r pipeline-table}
walkthrough <- safe_tar_read("vig_cr_pipeline_walkthrough")
if (!is.null(walkthrough)) {
  knitr::kable(walkthrough,
    caption = paste0("irishbuoys pipeline: ", nrow(walkthrough),
                     " plan files across data, QA, analysis, and output layers. ",
                     "Each plan file is a modular targets plan sourced by _targets.R. ",
                     "Source: plan_vignette_closeread.R."))
}
```

#### Simplified Flow

```{mermaid}
graph LR
  acq["ERDDAP fetch<br/>DuckDB store"]
  val["8 validation<br/>targets"]
  wave["61 wave<br/>analysis targets"]
  alert["Storm alerts<br/>+ email"]
  site["Dashboard<br/>+ pkgdown"]
  qa["QA gates<br/>Bronze/Silver/Gold"]
  acq --> val --> wave
  wave --> alert
  wave --> site
  qa --> site
```

::: {.card-footer}
Simplified view: <a href="https://erddap.marine.ie/" target="_blank">ERDDAP</a> fetch → 8 validation targets → 61 wave analysis targets → storm alerts + dashboard + <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a>. QA gates score Bronze/Silver/Gold before merge.
:::


# FOCUS Framework

## Row

### {.tabset}

#### FOCUS Diagram

```{mermaid}
flowchart LR
  F["Find<br/>Glob, Grep, Explore agent"]
  O["Organize<br/>AGENTS.md, Plan agent"]
  C["Condense<br/>Memory, ctx.yaml cache"]
  U["Understand<br/>r-debugger, critic"]
  Sy["Synthesize<br/>fixer, targets-runner"]
  F --> O --> C --> U --> Sy
```

::: {.card-footer}
FOCUS: how <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> navigates a project. Find → Organize → Condense → Understand → Synthesize. Each phase maps to specific agents, rules, and skills.
:::

#### Find

**Find** — locate relevant code, files, and patterns.

| Tool | Purpose |
|------|---------|
| `Glob` | Fast file pattern matching (`**/*.R`) |
| `Grep` | Content search with regex |
| `Explore` agent | Deep codebase exploration |
| `ctx.yaml` cache | Pre-indexed package signatures |

#### Organize

**Organize** — structure work into plans and tasks.

| Component | Role |
|-----------|------|
| <a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> | Entry point — indexes all infrastructure |
| Plan agent | Design implementation strategies |
| `architecture-planning` rule | Mandatory brainstorming before coding |
| `requirements-spec` skill | MUST/SHOULD/MAY for complex tasks |

#### Condense

**Condense** — compress context, cache knowledge.

| Component | Role |
|-----------|------|
| Memory files | Persist learnings across conversations |
| `context_survival.sh` | Save state before compaction |
| `ctx-yaml-cache` rule | Pre-compute package context |
| CHANGELOG.md | Record completed work, failed approaches |

#### Understand

**Understand** — debug, review, validate.

| Component | Role |
|-----------|------|
| `r-debugger` agent | Hypothesis-Experiment-Conclusion loops |
| `critic` agent | Read-only adversarial review |
| `quality-gates` skill | Bronze/Silver/Gold scoring |
| `systematic-debugging` rule | Scientific method for bugs |

#### Synthesize

**Synthesize** — produce outputs.

| Component | Role |
|-----------|------|
| `fixer` agent | Apply fixes from critic reports |
| `targets-runner` agent | Execute pipelines |
| `orchestrator-protocol` | Verify → review → fix → score loop |
| `r-package-workflow` | 9-step PR workflow |


# Build Info

## Row

```{r build-info}
#| results: asis
info <- safe_tar_read("vig_cr_build_info")
if (!is.null(info)) cat(info) else cat("*Run `tar_make()` locally to generate build info.*")
```
