---
title: "From Nix Shells to Project Outputs"
format:
  html:
    toc: false
    page-layout: full
    code-fold: true
    embed-resources: false
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

safe_tar_read <- function(name) {
  store <- tryCatch({
    owd <- setwd(here::here())
    on.exit(setwd(owd))
    normalizePath(targets::tar_config_get("store"))
  }, error = function(e) NULL)

  if (!is.null(store)) {
    obj <- tryCatch(targets::tar_read_raw(name, store = store), error = function(e) NULL)
    if (!is.null(obj)) return(obj)
  }

  for (d in c("../inst/extdata/vignettes", "inst/extdata/vignettes",
              file.path(here::here(), "inst/extdata/vignettes"))) {
    rds <- file.path(d, paste0(name, ".rds"))
    if (file.exists(rds)) return(readRDS(rds))
  }
  NULL
}

counts <- safe_tar_read("vig_cr_infra_counts")
if (is.null(counts)) {
  # Count files directly (reproducible, always current)
  claude_dir <- path.expand("~/.claude")

  n_rules <- length(list.files(file.path(claude_dir, "rules"), pattern = "\\.md$", recursive = TRUE))
  n_skills <- length(list.dirs(file.path(claude_dir, "skills"), recursive = FALSE))
  n_agents <- length(list.files(file.path(claude_dir, "agents"), pattern = "\\.md$", recursive = TRUE))
  n_hooks <- length(list.files(file.path(claude_dir, "hooks"), pattern = "\\.sh$"))
  n_commands <- length(list.files(file.path(claude_dir, "commands"), pattern = "\\.md$", recursive = TRUE))
  n_memory <- length(list.files(file.path(claude_dir, "memory"), pattern = "\\.md$", recursive = TRUE))

  counts <- list(
    n_rules = n_rules,
    n_skills = n_skills,
    n_agents = n_agents,
    n_hooks = n_hooks,
    n_commands = n_commands,
    n_memory = n_memory
  )
}
```


## Software Layers

```{mermaid}
%%| fig-cap: "Software stack from OS through [Nix](https://nixos.org/){target='_blank'} to project outputs. [rix](https://docs.ropensci.org/rix/){target='_blank'} translates `default.R` into `default.nix`. [targets](https://docs.ropensci.org/targets/){target='_blank'}, [crew](https://wlandau.github.io/crew/){target='_blank'}, and [mirai](https://cran.r-project.org/package=mirai){target='_blank'} orchestrate computation."
graph LR
  OS["OS: macOS / Linux"] --> Nix["Nix + rix"] --> Pkgs["R / Python<br/>packages"]
  Pkgs --> Orch["targets + crew<br/>+ mirai"]
  Orch --> LLM["LLM harness"]
  LLM --> Out["Outputs"]
  click OS "https://support.apple.com/macos" _blank
  click Nix "https://nixos.org/" _blank
  click Pkgs "https://docs.ropensci.org/rix/" _blank
  click Orch "https://docs.ropensci.org/targets/" _blank
  click LLM "https://github.com/JohnGavin/llm/blob/main/AGENTS.md" _blank
  click Out "https://johngavin.github.io/llm/" _blank
```

::: {.panel-tabset}

### How It Works

::: {.callout-note}
## Nix + rix
<a href="https://nixos.org/" target="_blank">Nix</a> provides hermetic, reproducible package management. <a href="https://docs.ropensci.org/rix/" target="_blank">rix</a> translates R-flavoured declarations in `default.R` into `default.nix` derivations. Every project gets its own isolated shell with pinned package versions.
:::

::: {.callout-tip}
## Orchestration
Above Nix sit the project-specific R and Python packages. <a href="https://docs.ropensci.org/targets/" target="_blank">targets</a> orchestrates pipelines, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> manages parallel workers, and <a href="https://cran.r-project.org/package=mirai" target="_blank">mirai</a> provides the async evaluation engine.
:::

::: {.callout-important}
## LLM Harness
At the top, `r counts$n_rules` rules, `r counts$n_skills` skills, `r counts$n_agents` agents, and `r counts$n_memory` memory files guide <a href="https://claude.com/claude-code" target="_blank">Claude Code</a> through finding, organising, and synthesising project work.
:::

:::


## LLM Harness

```{mermaid}
%%| fig-cap: "[AGENTS.md](https://github.com/JohnGavin/llm/blob/main/AGENTS.md){target='_blank'} is the entry point — indexing rules, skills, agents, hooks, commands, and memory files."
graph LR
  A["AGENTS.md<br/>Entry point"] --> R["Rules"]
  A --> S["Skills"]
  A --> G["Agents"]
  A --> H["Hooks"]
  A --> C["Commands"]
  A --> M["Memory"]
  click A "https://github.com/JohnGavin/llm/blob/main/AGENTS.md" _blank
  click R "https://github.com/JohnGavin/llm/tree/main/.claude/rules" _blank
  click S "https://github.com/JohnGavin/llm/tree/main/.claude/skills" _blank
  click G "https://github.com/JohnGavin/llm/tree/main/.claude/agents" _blank
  click H "https://github.com/JohnGavin/llm/tree/main/.claude/hooks" _blank
  click C "https://github.com/JohnGavin/llm/tree/main/.claude/commands" _blank
  click M "https://github.com/JohnGavin/llm/blob/main/AGENTS.md#memory-8-files" _blank
```

::: {.panel-tabset}

### Rules

```{r rules-table}
rules_dir <- file.path(path.expand("~/.claude"), "rules")
if (dir.exists(rules_dir)) {
  rule_files <- list.files(rules_dir, pattern = "\\.md$", full.names = TRUE, recursive = TRUE)

  # Extract rule name and first meaningful line as description
  rules_df <- lapply(rule_files, function(f) {
    lines <- readLines(f, warn = FALSE, n = 20)
    name <- tools::file_path_sans_ext(basename(f))

    # Find first non-empty, non-heading line for description
    desc_lines <- lines[!grepl("^#|^$|^---$|^Source:|^Consolidated", lines)]
    desc <- if (length(desc_lines) > 0) {
      trimws(gsub("\\*\\*|\\[|\\]\\([^)]+\\)", "", desc_lines[1]))
    } else {
      ""
    }
    # Truncate long descriptions
    if (nchar(desc) > 60) desc <- paste0(substr(desc, 1, 57), "...")

    data.frame(Rule = paste0("`", name, "`"), Description = desc, stringsAsFactors = FALSE)
  }) |> do.call(rbind, args = _)

  knitr::kable(rules_df, caption = paste0("All ", nrow(rules_df), " rules in ~/.claude/rules/"))
} else {
  cat("*Rules directory not found*")
}
```

### Skills

```{r skills-table}
skill_cats <- safe_tar_read("vig_cr_skill_categories")
if (!is.null(skill_cats)) {
  # Add example links for major categories
  skill_cats$Examples <- dplyr::case_when(
    skill_cats$category == "R Package Dev" ~
      '<a href="https://testthat.r-lib.org/" target="_blank">testthat</a>, <a href="https://lifecycle.r-lib.org/" target="_blank">lifecycle</a>, <a href="https://rlang.r-lib.org/" target="_blank">rlang</a>',
    skill_cats$category == "Data & Analysis" ~
      '<a href="https://rstudio.github.io/pointblank/" target="_blank">pointblank</a>, EDA, modelling',
    skill_cats$category == "Targets & Pipelines" ~
      'Pipeline spec, vignettes, <a href="https://wlandau.github.io/crew/" target="_blank">crew</a> ops',
    skill_cats$category == "Shiny & Web" ~
      '<a href="https://rstudio.github.io/bslib/" target="_blank">bslib</a>, async, <a href="https://posit-dev.github.io/r-shinylive/" target="_blank">Shinylive</a>',
    skill_cats$category == "Quarto & Docs" ~
      'Websites, dashboards, alt text',
    skill_cats$category == "DevOps & CI" ~
      'GitHub Actions, <a href="https://pkgdown.r-lib.org/" target="_blank">pkgdown</a>',
    TRUE ~ ""
  )

  skill_cats |>
    dplyr::select(Category = category, Count = count, Examples) |>
    knitr::kable(caption = paste0("All ", sum(skill_cats$count),
                                  " skills from ~/.claude/skills/MANIFEST.md"))
} else {
  cat("*Run `tar_make()` locally to generate skill counts*")
}
```

### Agents

```{r agents-table}
agents_dir <- file.path(path.expand("~/.claude"), "agents")
if (dir.exists(agents_dir)) {
  agent_files <- list.files(agents_dir, pattern = "\\.md$", full.names = TRUE, recursive = TRUE)

  agents_df <- lapply(agent_files, function(f) {
    lines <- readLines(f, warn = FALSE, n = 30)
    name <- tools::file_path_sans_ext(basename(f))

    # Try to extract model from frontmatter or content
    model_line <- grep("model:", lines, value = TRUE, ignore.case = TRUE)
    model <- if (length(model_line) > 0) {
      gsub(".*model:\\s*", "", model_line[1])
    } else {
      "sonnet"
    }

    # Find trigger/use-case description
    trigger_lines <- lines[grepl("Use when|Trigger|Debug|Run|Review|Fix|Build", lines, ignore.case = TRUE)]
    trigger <- if (length(trigger_lines) > 0) {
      desc <- trimws(gsub("^[#*-]+\\s*|\\*\\*|Use when:?\\s*", "", trigger_lines[1]))
      if (nchar(desc) > 50) paste0(substr(desc, 1, 47), "...") else desc
    } else {
      ""
    }

    data.frame(Agent = paste0("`", name, "`"), Model = model, Trigger = trigger, stringsAsFactors = FALSE)
  }) |> do.call(rbind, args = _)

  knitr::kable(agents_df, caption = paste0("All ", nrow(agents_df), " agents in ~/.claude/agents/"))
} else {
  cat("*Agents directory not found*")
}
```

### Hooks & Memory

**Hooks** — shell scripts triggered by events:

```{r hooks-table}
hooks_dir <- file.path(path.expand("~/.claude"), "hooks")
if (dir.exists(hooks_dir)) {
  hook_files <- list.files(hooks_dir, pattern = "\\.sh$", full.names = TRUE)

  hooks_df <- lapply(hook_files, function(f) {
    lines <- readLines(f, warn = FALSE, n = 15)
    name <- basename(f)

    # Extract event from filename pattern
    event <- dplyr::case_when(
      grepl("init", name) ~ "Session start",
      grepl("protection", name) ~ "Pre-edit",
      grepl("survival|compact", name) ~ "Pre-compact",
      grepl("monitor", name) ~ "Post-tool",
      grepl("stop", name) ~ "Session stop",
      grepl("health", name) ~ "Post-write",
      TRUE ~ "Hook"
    )

    # Extract purpose from first comment line
    comment_lines <- grep("^#[^!]", lines, value = TRUE)
    purpose <- if (length(comment_lines) > 0) {
      desc <- trimws(gsub("^#+\\s*", "", comment_lines[1]))
      if (nchar(desc) > 40) paste0(substr(desc, 1, 37), "...") else desc
    } else {
      ""
    }

    data.frame(Hook = paste0("`", name, "`"), Event = event, Purpose = purpose, stringsAsFactors = FALSE)
  }) |> do.call(rbind, args = _)

  knitr::kable(hooks_df, caption = paste0("All ", nrow(hooks_df), " hooks"))
} else {
  cat("*Hooks directory not found*")
}
```

**Memory** — file-based persistence across conversations:

```{r memory-table}
memory_dir <- file.path(path.expand("~/.claude"), "memory")
if (dir.exists(memory_dir)) {
  memory_files <- list.files(memory_dir, pattern = "\\.md$", full.names = TRUE, recursive = TRUE)

  memory_df <- lapply(memory_files, function(f) {
    lines <- readLines(f, warn = FALSE, n = 10)
    name <- tools::file_path_sans_ext(basename(f))

    # Extract first meaningful line as purpose
    desc_lines <- lines[!grepl("^#|^$|^---$", lines)]
    purpose <- if (length(desc_lines) > 0) {
      desc <- trimws(desc_lines[1])
      if (nchar(desc) > 50) paste0(substr(desc, 1, 47), "...") else desc
    } else {
      ""
    }

    data.frame(File = paste0("`", name, "`"), Purpose = purpose, stringsAsFactors = FALSE)
  }) |> do.call(rbind, args = _)

  knitr::kable(memory_df, caption = paste0("All ", nrow(memory_df), " memory files"))
} else {
  cat("*Memory directory not found*")
}
```

:::


## irishbuoys: A Real Project Pipeline

The <a href="https://github.com/JohnGavin/irishbuoys" target="_blank">irishbuoys</a> R package is an independent project that monitors wave conditions at Marine Institute buoy stations around Ireland. Its <a href="https://johngavin.github.io/irishbuoys/" target="_blank">pkgdown site</a> is deployed separately. The pipeline below illustrates how the software layers described above come together in practice — from <a href="https://erddap.marine.ie/" target="_blank">ERDDAP</a> data acquisition through QA validation to multiple output formats.

```{mermaid}
%%| fig-cap: "[irishbuoys](https://github.com/JohnGavin/irishbuoys){target='_blank'} pipeline ([pkgdown site](https://johngavin.github.io/irishbuoys/){target='_blank'}). 19 plan files, ~294 [targets](https://docs.ropensci.org/targets/){target='_blank'}. Data from [ERDDAP](https://erddap.marine.ie/){target='_blank'}, stored in [DuckDB](https://duckdb.org/){target='_blank'}."
graph LR
  acq["ERDDAP<br/>API"] --> val["Validation<br/>8 targets"] --> qc["QC<br/>bounds"]
  qc --> wave["Wave Analysis<br/>61 targets"]
  wave --> joint["Joint<br/>Analysis"] --> sev["Spatial<br/>Extremes"]
  wave --> pred["Predictions"] --> alert["Storm<br/>Alerts"] --> email["Email<br/>blastula"]
  wave --> dash["Dashboard<br/>bslib"]
  wave --> api["Static<br/>API"]
  joint --> site["pkgdown<br/>Site"]
  qa["QA Gates<br/>Bronze/Silver/Gold"] --> site
  click acq "https://erddap.marine.ie/" _blank
  click val "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_data_validation.R" _blank
  click qc "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_quality_control.R" _blank
  click wave "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_wave_analysis.R" _blank
  click joint "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_joint_analysis.R" _blank
  click sev "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_spatial_extreme_values.R" _blank
  click pred "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_predictions.R" _blank
  click alert "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_storm_alert.R" _blank
  click email "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_email_report.R" _blank
  click dash "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_dashboard.R" _blank
  click api "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_api.R" _blank
  click site "https://johngavin.github.io/irishbuoys/" _blank
  click qa "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_qa_gates.R" _blank
```

**Simplified flow:**

```{mermaid}
%%| fig-cap: "Simplified: [ERDDAP](https://erddap.marine.ie/){target='_blank'} → validation → wave analysis → outputs. [QA gates](https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_qa_gates.R){target='_blank'} score Bronze/Silver/Gold before merge."
graph LR
  acq["ERDDAP"] --> val["8 validation"] --> wave["61 analysis"]
  wave --> alert["Alerts + email"]
  wave --> site["Dashboard + pkgdown"]
  qa["QA gates"] --> site
  click acq "https://erddap.marine.ie/" _blank
  click val "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_data_validation.R" _blank
  click wave "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_wave_analysis.R" _blank
  click alert "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_storm_alert.R" _blank
  click site "https://johngavin.github.io/irishbuoys/" _blank
  click qa "https://github.com/JohnGavin/irishbuoys/blob/main/R/tar_plans/plan_qa_gates.R" _blank
```

::: {.panel-tabset}

### Plan Files

```{r pipeline-table}
walkthrough <- safe_tar_read("vig_cr_pipeline_walkthrough")
if (!is.null(walkthrough)) {
  knitr::kable(walkthrough,
    caption = paste0("irishbuoys pipeline: ", nrow(walkthrough),
                     " plan files across data, QA, analysis, and output layers."))
}
```

:::


## FOCUS Framework

```{mermaid}
%%| fig-cap: "FOCUS: how [Claude Code](https://claude.com/claude-code){target='_blank'} navigates a project. Each phase maps to specific agents, rules, and skills."
flowchart LR
  F["Find<br/>Glob, Grep, Explore"] --> O["Organize<br/>AGENTS.md, Plan"]
  O --> C["Condense<br/>Memory, ctx.yaml"]
  C --> U["Understand<br/>r-debugger, critic"]
  U --> Sy["Synthesize<br/>fixer, targets-runner"]
  click F "https://github.com/JohnGavin/llm/tree/main/.claude/agents" _blank
  click O "https://github.com/JohnGavin/llm/blob/main/AGENTS.md" _blank
  click C "https://github.com/JohnGavin/llm/blob/main/AGENTS.md#memory-8-files" _blank
  click U "https://github.com/JohnGavin/llm/tree/main/.claude/agents" _blank
  click Sy "https://github.com/JohnGavin/llm/tree/main/.claude/agents" _blank
```

::: {.panel-tabset}

### Find

| Tool | Purpose |
|------|---------|
| `Glob` | Fast file pattern matching (`**/*.R`) |
| `Grep` | Content search with regex |
| `Explore` agent | Deep codebase exploration |
| `ctx.yaml` cache | Pre-indexed package signatures |

### Organize

| Component | Role |
|-----------|------|
| <a href="https://github.com/JohnGavin/llm/blob/main/AGENTS.md" target="_blank">AGENTS.md</a> | Entry point — indexes all infrastructure |
| Plan agent | Design implementation strategies |
| `architecture-planning` rule | Mandatory brainstorming before coding |

### Condense

| Component | Role |
|-----------|------|
| Memory files | Persist learnings across conversations |
| `context_survival.sh` | Save state before compaction |
| CHANGELOG.md | Record completed work, failed approaches |

### Understand

| Component | Role |
|-----------|------|
| `r-debugger` agent | Hypothesis-Experiment-Conclusion loops |
| `critic` agent | Read-only adversarial review |
| `quality-gates` skill | Bronze/Silver/Gold scoring |

### Synthesize

| Component | Role |
|-----------|------|
| `fixer` agent | Apply fixes from critic reports |
| `targets-runner` agent | Execute pipelines |
| `orchestrator-protocol` | Verify → review → fix → score loop |

:::


## Build Info

```{r build-info}
#| results: asis
info <- safe_tar_read("vig_cr_build_info")
if (!is.null(info)) cat(info) else cat("*Run `tar_make()` locally to generate build info.*")
```
