---
title: "Global Config Evolution Dashboard"
format:
  html:
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
---

{{< include /_includes/hover-popups.qmd >}}

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(scales)
library(gt)

# Theme for dark mode compatibility
theme_dashboard <- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold")
  )

# Try to load from local pulse data, otherwise count files directly
config_dir <- path.expand("~/.claude/logs/config")
parquet_files <- list.files(config_dir, pattern = "\\.parquet$", full.names = TRUE)

config_data <- NULL
if (length(parquet_files) > 0 && requireNamespace("DBI", quietly = TRUE) && requireNamespace("duckdb", quietly = TRUE)) {
  tryCatch({
    con <- DBI::dbConnect(duckdb::duckdb())
    config_data <- DBI::dbGetQuery(con, sprintf(
      "SELECT * FROM read_parquet('%s/*.parquet')",
      config_dir
    )) |>
      mutate(
        snapshot_date = as.Date(snapshot_date),
        value = as.numeric(value)
      )
    DBI::dbDisconnect(con, shutdown = TRUE)
  }, error = function(e) NULL)
}

# Fallback: count files directly (works in CI).
# In CI, ~/.claude doesn't exist — the repo-committed .claude/ does (per #144).
# Probe in priority order so this also works locally where ~/.claude has the
# same content via symlinks.
if (is.null(config_data)) {
  # Probe repo-relative path first (works in CI and locally via symlink).
  # The user-level ~/.claude path is excluded: vignettes must be reproducible
  # from the checked-in source tree alone, not from reviewer-local state.
  .candidates <- c(here::here(".claude"))
  claude_dir <- Find(dir.exists, .candidates)
  if (is.null(claude_dir)) claude_dir <- here::here(".claude")  # last-resort string

  # Count rules
  rules_dir <- file.path(claude_dir, "rules")
  if (dir.exists(rules_dir)) {
    rule_files <- list.files(rules_dir, pattern = "\\.md$", full.names = TRUE, recursive = TRUE)
    n_rules <- length(rule_files)
    rule_lines <- sapply(rule_files, function(f) length(readLines(f, warn = FALSE)))
  } else {
    n_rules <- 0
    rule_lines <- numeric(0)
  }

 # Count skills
  skills_dir <- file.path(claude_dir, "skills")
  if (dir.exists(skills_dir)) {
    skill_dirs <- list.dirs(skills_dir, recursive = FALSE)
    n_skills <- length(skill_dirs)
    skill_files <- list.files(skills_dir, pattern = "SKILL\\.md$", full.names = TRUE, recursive = TRUE)
    skill_lines <- sapply(skill_files, function(f) length(readLines(f, warn = FALSE)))
  } else {
    n_skills <- 0
    skill_lines <- numeric(0)
  }

  # Count hooks
  hooks_dir <- file.path(claude_dir, "hooks")
  if (dir.exists(hooks_dir)) {
    hook_files <- list.files(hooks_dir, pattern = "\\.sh$", full.names = TRUE)
    n_hooks <- length(hook_files)
    hook_lines <- sapply(hook_files, function(f) length(readLines(f, warn = FALSE)))
  } else {
    n_hooks <- 0
    hook_lines <- numeric(0)
  }

  # Estimate total tokens
  all_files <- c(
    list.files(claude_dir, pattern = "\\.md$", full.names = TRUE, recursive = TRUE),
    list.files(claude_dir, pattern = "\\.sh$", full.names = TRUE, recursive = TRUE)
  )
  all_files <- all_files[!grepl("/logs/", all_files)]
  total_bytes <- sum(file.info(all_files)$size, na.rm = TRUE)
  total_tokens <- round(total_bytes / 4)

  # Create summary
  fallback_counts <- list(
    n_rules = n_rules,
    n_skills = n_skills,
    n_hooks = n_hooks,
    total_tokens = total_tokens,
    rule_lines = rule_lines,
    skill_lines = skill_lines,
    hook_lines = hook_lines,
    total_rule_lines = sum(rule_lines),
    total_skill_lines = sum(skill_lines),
    total_hook_lines = sum(hook_lines),
    rules_over_150 = sum(rule_lines > 150),
    avg_rule_lines = if (n_rules > 0) round(mean(rule_lines)) else 0,
    avg_skill_lines = if (n_skills > 0) round(mean(skill_lines)) else 0
  )
}

# Create latest dataset if config_data exists (used by all metric chunks)
latest <- NULL
if (!is.null(config_data)) {
  latest <- config_data |> filter(snapshot_date == max(snapshot_date))
}
```

Track the evolution of Claude Code global configuration: rules, skills, agents, hooks, and memories.

## Current Snapshot

::: {.grid}

::: {.g-col-3}
### Rules
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "rules", item == "all", metric == "count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_counts$n_rules)
}
```
:::

::: {.g-col-3}
### Skills
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "skills", item == "all", metric == "count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_counts$n_skills)
}
```
:::

::: {.g-col-3}
### Hooks
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "hooks", item == "all", metric == "count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_counts$n_hooks)
}
```
:::

::: {.g-col-3}
### Total Tokens
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "total", metric == "tokens_estimate") |> pull(value)
  cat("####", comma(n))
} else {
  cat("####", comma(fallback_counts$total_tokens))
}
```
:::

:::

::: {.panel-tabset}

## Size Distribution

```{r size-distribution}
#| fig-width: 10
#| fig-height: 6
#| fig-alt: "Bar chart showing line counts by category: rules, skills, hooks"

if (!is.null(latest)) {
  latest |>
    filter(
      item == "all",
      metric == "total_lines",
      category %in% c("rules", "skills", "hooks", "scripts", "memory")
    ) |>
    mutate(category = factor(category, levels = c("skills", "rules", "scripts", "hooks", "memory"))) |>
    ggplot(aes(x = category, y = value, fill = category)) +
    geom_col() +
    geom_text(aes(label = comma(value)), vjust = -0.5, size = 4) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    scale_fill_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Config Size by Category",
      subtitle = paste("Snapshot:", max(config_data$snapshot_date)),
      x = NULL,
      y = "Lines of Code"
    ) +
    theme_dashboard +
    theme(legend.position = "none")
} else {
  # Fallback chart
  tibble(
    category = c("skills", "rules", "hooks"),
    value = c(fallback_counts$total_skill_lines, fallback_counts$total_rule_lines, fallback_counts$total_hook_lines)
  ) |>
    mutate(category = factor(category, levels = c("skills", "rules", "hooks"))) |>
    ggplot(aes(x = category, y = value, fill = category)) +
    geom_col() +
    geom_text(aes(label = comma(value)), vjust = -0.5, size = 4) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    scale_fill_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Config Size by Category",
      subtitle = paste("Live count:", Sys.Date()),
      x = NULL,
      y = "Lines of Code"
    ) +
    theme_dashboard +
    theme(legend.position = "none")
}
```

## Top Rules

```{r top-rules}
if (!is.null(latest)) {
  latest |>
    filter(category == "rules", metric == "lines", item != "all") |>
    arrange(desc(value)) |>
    head(15) |>
    mutate(
      status = case_when(
        value > 150 ~ "Over threshold",
        value > 100 ~ "Warning",
        TRUE ~ "OK"
      )
    ) |>
    gt() |>
    cols_label(
      item = "Rule",
      value = "Lines",
      status = "Status"
    ) |>
    cols_hide(c(snapshot_date, category, metric)) |>
    data_color(
      columns = value,
      palette = c("#69d4a0", "#ffd93d", "#f08080"),
      domain = c(0, 200)
    ) |>
    tab_header(
      title = "Largest Rules",
      subtitle = "Threshold: 150 lines (per CLAUDE.md)"
    )
} else if (length(fallback_counts$rule_lines) > 0) {
  # Fallback table from direct file counts

  tibble(
    rule = basename(names(fallback_counts$rule_lines)) |> tools::file_path_sans_ext(),
    lines = as.integer(fallback_counts$rule_lines)
  ) |>
    arrange(desc(lines)) |>
    head(15) |>
    mutate(
      status = case_when(
        lines > 150 ~ "Over threshold",
        lines > 100 ~ "Warning",
        TRUE ~ "OK"
      )
    ) |>
    gt() |>
    cols_label(rule = "Rule", lines = "Lines", status = "Status") |>
    data_color(
      columns = lines,
      palette = c("#69d4a0", "#ffd93d", "#f08080"),
      domain = c(0, 200)
    ) |>
    tab_header(
      title = "Largest Rules",
      subtitle = "Threshold: 150 lines"
    )
} else {
  cat("*No rule data available in CI environment.*")
}
```

## Top Skills

```{r top-skills}
if (!is.null(latest)) {
  latest |>
    filter(category == "skills", metric == "lines", item != "all") |>
    arrange(desc(value)) |>
    head(15) |>
    gt() |>
    cols_label(
      item = "Skill",
      value = "Lines"
    ) |>
    cols_hide(c(snapshot_date, category, metric)) |>
    data_color(
      columns = value,
      palette = "Blues"
    ) |>
    tab_header(
      title = "Largest Skills"
    )
} else if (length(fallback_counts$skill_lines) > 0) {
  tibble(
    skill = basename(dirname(names(fallback_counts$skill_lines))),
    lines = as.integer(fallback_counts$skill_lines)
  ) |>
    arrange(desc(lines)) |>
    head(15) |>
    gt() |>
    cols_label(skill = "Skill", lines = "Lines") |>
    data_color(columns = lines, palette = "Blues") |>
    tab_header(title = "Largest Skills")
} else {
  cat("*No skill data available in CI environment.*")
}
```

## Quality Metrics

```{r quality-table}
if (!is.null(latest)) {
  quality_metrics <- latest |>
    filter(item == "all") |>
    filter(
      (category == "rules" & metric %in% c("count", "avg_lines", "over_150_lines", "with_frontmatter")) |
      (category == "skills" & metric %in% c("count", "avg_lines")) |
      (category == "total" & metric == "tokens_estimate")
    ) |>
    mutate(
      label = case_when(
        category == "rules" & metric == "count" ~ "Total Rules",
        category == "rules" & metric == "avg_lines" ~ "Avg Rule Size (lines)",
        category == "rules" & metric == "over_150_lines" ~ "Rules Over Threshold",
        category == "rules" & metric == "with_frontmatter" ~ "Rules with Frontmatter",
        category == "skills" & metric == "count" ~ "Total Skills",
        category == "skills" & metric == "avg_lines" ~ "Avg Skill Size (lines)",
        category == "total" & metric == "tokens_estimate" ~ "Total Config Tokens",
        TRUE ~ paste(category, metric)
      )
    ) |>
    select(label, value)

  quality_metrics |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE)
} else {
  tibble(
    label = c("Total Rules", "Avg Rule Size (lines)", "Rules Over Threshold",
              "Total Skills", "Avg Skill Size (lines)", "Total Config Tokens"),
    value = c(fallback_counts$n_rules, fallback_counts$avg_rule_lines, fallback_counts$rules_over_150,
              fallback_counts$n_skills, fallback_counts$avg_skill_lines, fallback_counts$total_tokens)
  ) |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE)
}
```

## Evolution

```{r evolution-chart}
#| fig-width: 10
#| fig-height: 5
#| fig-alt: "Line chart showing config evolution over time"

if (!is.null(latest) && n_distinct(config_data$snapshot_date) > 1) {
  config_data |>
    filter(
      item == "all",
      metric == "count",
      category %in% c("rules", "skills", "hooks")
    ) |>
    ggplot(aes(x = snapshot_date, y = value, color = category)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Config Growth Over Time",
      x = NULL,
      y = "Count",
      color = "Category"
    ) +
    theme_dashboard
} else {
  cat("*Run `~/.claude/scripts/config_pulse.sh` daily to collect time-series data.*

Current counts: ", fallback_counts$n_rules, " rules, ", fallback_counts$n_skills, " skills, ", fallback_counts$n_hooks, " hooks")
}
```

:::

---

**Data Collection:** Run `~/.claude/scripts/config_pulse.sh` daily to collect snapshots stored in `~/.claude/logs/config/`.

*Last updated: `r Sys.time()`*

## Architecture & Intent

Every `.claude/` file is a unit of knowledge: a rule encodes a constraint, a
skill encodes a capability, a hook encodes an enforcement mechanism. This
dashboard treats that knowledge base as data — tracking counts, sizes, and
growth rates across sessions. The dashboard is the *pull* view (open it and
browse); the daily [config-change digest](#297) is the *push* view (it arrives
in your inbox each morning with a summary of what changed and why).

### Pipeline overview

The diagram below shows the two complementary flows. The *pulse* flow (top)
powers this dashboard. The *digest* flow (bottom) powers the morning email.
Both draw from the same source of truth: the `.claude/` directory tree under
version control.

::: {aria-label="Pipeline diagram showing two flows. Top: config_pulse.sh writes daily Parquet snapshots to logs/config/; the config-evolution dashboard reads those snapshots via DuckDB. Bottom: git history feeds config_change_digest.R; that R script produces a markdown digest; send_config_digest_email.R sends the digest via blastula and Gmail credentials; the email arrives in the inbox. A separate overnight chain shows self-review-stage1 at 02:30 running the overnight Claude review; self-review-verify at 08:00 running the email dispatch; and a scheduled-task health email cross-linking to issue 300."}

```{mermaid}
flowchart LR
    A[".claude/ tree\n(git history)"] --> B["config_pulse.sh\n(daily, launchd)"]
    B --> C["logs/config/*.parquet\n(DuckDB snapshots)"]
    C --> D["config-evolution\ndashboard\n(this page)"]

    A --> E["config_change_digest.R\n(--since 24h)"]
    E --> F["markdown digest\n(themes + lessons)"]
    F --> G["send_config_digest_email.R\n(blastula + Gmail)"]
    G --> H["inbox\n(08:05 daily)"]

    A --> I["self-review-stage1\n(02:30, launchd)"]
    I --> J["stage1 findings\n(unified.duckdb)"]
    J --> K["self-review-verify\n(08:00, launchd)"]
    K --> G

    click B "https://github.com/JohnGavin/llm/blob/main/.claude/scripts/config_pulse.sh#L15" "config_pulse.sh — writes daily Parquet to logs/config/"
    click C "https://github.com/JohnGavin/llm/blob/main/.claude/scripts/config_pulse.sh#L168" "config_pulse.sh — DuckDB COPY converts CSV to Parquet"
    click E "https://github.com/JohnGavin/llm/blob/main/.claude/scripts/config_change_digest.R#L102" "config_change_digest.R — git log --numstat grouped by category"
    click G "https://github.com/JohnGavin/llm/blob/main/.claude/scripts/send_config_digest_email.R#L386" "send_config_digest_email.R — smtp_send via Gmail creds"
    click I "https://github.com/JohnGavin/llm/blob/main/.claude/launchd/com.claude.self-review-stage1.plist#L1" "self-review-stage1 launchd plist (02:30)"
    click K "https://github.com/JohnGavin/llm/blob/main/.claude/launchd/com.claude.self-review-verify.plist#L1" "self-review-verify launchd plist (08:00)"
```

:::

### Key ideas for a new reader

- **Git history is the source of truth.** Every count, size, and delta in this
  dashboard and in the digest email is derived from `git log` or `list.files()`
  on the checked-in `.claude/` tree. Nothing is manually entered or hardcoded
  ([`dynamic-prose-values`](https://github.com/JohnGavin/llm/blob/main/.claude/rules/dynamic-prose-values.md#L1) rule).

- **Local compute, cloud delivery.** The pulse script and the digest aggregator
  run on the local machine (inside the Nix shell). Gmail credentials
  (`GMAIL_USERNAME`, `GMAIL_APP_PASSWORD`) live in the cloud
  ([llmtelemetry](https://github.com/JohnGavin/llmtelemetry) CI environment);
  the local jobs dispatch a `gh workflow run` so that the blastula send happens
  where the credentials are.

- **Jobs are idempotent and self-healing.** Each launchd plist has a lock file
  and a bounded timeout. If a run is skipped (laptop lid closed), the next run
  picks up from `--since <last-run-timestamp>`. Missing one day does not corrupt
  the time-series.

- **Backups live in a different failure domain.** The parquet snapshots and the
  unified.duckdb telemetry database are git-tracked or backed up to a separate
  path. If the local machine is lost, the git history of `.claude/` is the
  recovery path ([`backup-architecture`](https://github.com/JohnGavin/llm/blob/main/.claude/rules/backup-architecture.md#L1) rule).

- **The verify step is a safety net.** The 08:00 `self-review-verify` job
  ([llm #235](https://github.com/JohnGavin/llm/issues/235)) re-checks that the
  02:30 stage-1 job completed successfully before dispatching the email. If
  stage 1 failed silently, the verify email says so.

### Questions a new reader should ask

- *Where does this number come from?* Each panel traces to either
  `.claude/logs/config/*.parquet` (pulse path) or a live `list.files()` call
  (fallback path). The "Data sources" section in Methodology lists every source.

- *Is it reproducible or captured?* The count of rules today is reproducible
  (re-run `list.files()` on the same commit). The count of rules *three months
  ago* requires the parquet snapshots — those are captured and non-reproducible
  if the snapshots are lost.

- *What happens if the overnight job does not run?* The 08:00 verify job
  detects the gap and sends an alert rather than a normal digest. The verify
  email is the observable signal of pipeline health; see
  [#235](https://github.com/JohnGavin/llm/issues/235) and the
  [scheduled-task health email (#300)](https://github.com/JohnGavin/llm/issues/300).

- *How do I add a rule/skill and see it reflected here?* Add the file to the
  appropriate `.claude/` subdirectory and commit. The next time `config_pulse.sh`
  runs (daily, 08:05 via launchd), a new row appears in the parquet snapshot and
  this dashboard's counts update on the next render.

### How to read each panel and the digest email together

| Dashboard panel | What it shows | Digest email equivalent |
|---|---|---|
| Current Snapshot | Live counts of rules, skills, hooks, total tokens | "Category summary" table (files changed × lines ±) |
| Size Distribution | Total lines per category at latest snapshot | Implicit in "themes today" grouping |
| Top Rules / Top Skills | Largest individual files | Named in "lessons learnt" per theme |
| Quality Metrics | Rules over 150-line threshold, average sizes | Flagged as "over-threshold" items in digest |
| Evolution | Count growth over time (requires ≥2 snapshots) | Not in digest — trend context only |

The digest email is produced by
[`config_change_digest.R`](https://github.com/JohnGavin/llm/blob/main/.claude/scripts/config_change_digest.R#L102)
and sent by
[`send_config_digest_email.R`](https://github.com/JohnGavin/llm/blob/main/.claude/scripts/send_config_digest_email.R#L386)
via the cron wrapper
[`bin/config_digest_cron.sh`](https://github.com/JohnGavin/llm/blob/main/bin/config_digest_cron.sh#L6).
The email bridge pattern (local compute → `gh workflow run` → llmtelemetry CI
→ blastula/Gmail) is documented in
[llm PR #295](https://github.com/JohnGavin/llm/pull/295) and the companion
[llmtelemetry PR #211](https://github.com/JohnGavin/llmtelemetry/pull/211).

For the scrollytelling walkthrough of the same config data (five progressive
scenes: inventory → categories → age → hotspots → distributions), see the
companion vignette
[`vignettes/articles/scrolly-config-evolution.qmd`](articles/scrolly-config-evolution.qmd).

## Methodology

### What this vignette computes

This vignette tracks the evolution of the Claude Code global configuration — counts and line sizes of rules, skills, and hooks — over time. It reads time-series snapshots from `~/.claude/logs/config/*.parquet` via DuckDB when available, falling back to live `list.files()` counts from the `.claude/` directory tree when no snapshot data exists (CI environment). Key metrics computed include total file counts, average sizes, files over the 150-line threshold, and estimated token usage.

### Data sources

- `~/.claude/logs/config/*.parquet`: daily snapshots produced by `~/.claude/scripts/config_pulse.sh`; each row is one metric for one category (rules/skills/hooks) at one snapshot date
- Live fallback: `list.files()` on `.claude/rules/`, `.claude/skills/`, `.claude/hooks/` directories — used in CI where no parquet snapshots exist
- Token estimate: total bytes of all `.md` and `.sh` files under `.claude/` divided by 4

### AI disclosure

This vignette was developed with assistance from Anthropic's Claude (model: Opus 4.7 and Sonnet 4.6). AI helped with code structure, prose drafting, and visualization choices. All analytical decisions and data interpretations are the author's responsibility.

{{< include ../_includes/qr-footer.qmd >}}
