---
title: "Global Config Evolution Dashboard"
format:
  html:
    page-layout: full
    toc: true
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(scales)
library(gt)
library(DBI)
library(duckdb)

# Theme for dark mode compatibility
theme_dashboard <- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold")
  )

# Load config pulse data
config_dir <- path.expand("~/.claude/logs/config")
parquet_files <- list.files(config_dir, pattern = "\\.parquet$", full.names = TRUE)

if (length(parquet_files) > 0) {
  con <- dbConnect(duckdb())
  on.exit(dbDisconnect(con, shutdown = TRUE))

  # Read all parquet files
  config_data <- dbGetQuery(con, sprintf(
    "SELECT * FROM read_parquet('%s/*.parquet')",
    config_dir
  )) |>
    mutate(
      snapshot_date = as.Date(snapshot_date),
      value = as.numeric(value)
    )
} else {
  config_data <- NULL
}
```

Track the evolution of Claude Code global configuration: rules, skills, agents, hooks, and memories.

## Current Snapshot

::: {.grid}

::: {.g-col-3}
### Rules
```{r}
#| results: asis
if (!is.null(config_data)) {
  latest <- config_data |> filter(snapshot_date == max(snapshot_date))
  n_rules <- latest |> filter(category == "rules", item == "all", metric == "count") |> pull(value)
  cat("####", n_rules)
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-3}
### Skills
```{r}
#| results: asis
if (!is.null(config_data)) {
  n_skills <- latest |> filter(category == "skills", item == "all", metric == "count") |> pull(value)
  cat("####", n_skills)
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-3}
### Hooks
```{r}
#| results: asis
if (!is.null(config_data)) {
  n_hooks <- latest |> filter(category == "hooks", item == "all", metric == "count") |> pull(value)
  cat("####", n_hooks)
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-3}
### Total Tokens
```{r}
#| results: asis
if (!is.null(config_data)) {
  total_tokens <- latest |> filter(category == "total", metric == "tokens_estimate") |> pull(value)
  cat("####", comma(total_tokens))
} else {
  cat("#### No data")
}
```
:::

:::

## Size Distribution

```{r size-distribution}
#| fig-width: 10
#| fig-height: 6
#| fig-alt: "Stacked bar chart showing line counts by category: rules, skills, hooks, scripts"

if (!is.null(config_data)) {
  latest |>
    filter(
      item == "all",
      metric == "total_lines",
      category %in% c("rules", "skills", "hooks", "scripts", "memory")
    ) |>
    mutate(category = factor(category, levels = c("skills", "rules", "scripts", "hooks", "memory"))) |>
    ggplot(aes(x = category, y = value, fill = category)) +
    geom_col() +
    geom_text(aes(label = comma(value)), vjust = -0.5, size = 4) +
    scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
    scale_fill_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Config Size by Category",
      subtitle = paste("Snapshot:", max(config_data$snapshot_date)),
      x = NULL,
      y = "Lines of Code"
    ) +
    theme_dashboard +
    theme(legend.position = "none")
}
```

## Top Rules by Size

```{r top-rules}
if (!is.null(config_data)) {
  latest |>
    filter(category == "rules", metric == "lines", item != "all") |>
    arrange(desc(value)) |>
    head(15) |>
    mutate(
      status = case_when(
        value > 150 ~ "Over threshold",
        value > 100 ~ "Warning",
        TRUE ~ "OK"
      )
    ) |>
    gt() |>
    cols_label(
      item = "Rule",
      value = "Lines",
      status = "Status"
    ) |>
    cols_hide(c(snapshot_date, category, metric)) |>
    data_color(
      columns = value,
      palette = c("#69d4a0", "#ffd93d", "#f08080"),
      domain = c(0, 200)
    ) |>
    tab_header(
      title = "Largest Rules",
      subtitle = "Threshold: 150 lines (per CLAUDE.md)"
    )
}
```

## Top Skills by Size

```{r top-skills}
if (!is.null(config_data)) {
  latest |>
    filter(category == "skills", metric == "lines", item != "all") |>
    arrange(desc(value)) |>
    head(15) |>
    gt() |>
    cols_label(
      item = "Skill",
      value = "Lines"
    ) |>
    cols_hide(c(snapshot_date, category, metric)) |>
    data_color(
      columns = value,
      palette = "Blues"
    ) |>
    tab_header(
      title = "Largest Skills"
    )
}
```

## Quality Metrics

```{r quality-table}
if (!is.null(config_data)) {
  quality_metrics <- latest |>
    filter(item == "all") |>
    filter(
      (category == "rules" & metric %in% c("count", "avg_lines", "over_150_lines", "with_frontmatter")) |
      (category == "skills" & metric %in% c("count", "avg_lines")) |
      (category == "total" & metric == "tokens_estimate")
    ) |>
    mutate(
      label = case_when(
        category == "rules" & metric == "count" ~ "Total Rules",
        category == "rules" & metric == "avg_lines" ~ "Avg Rule Size (lines)",
        category == "rules" & metric == "over_150_lines" ~ "Rules Over Threshold",
        category == "rules" & metric == "with_frontmatter" ~ "Rules with Frontmatter",
        category == "skills" & metric == "count" ~ "Total Skills",
        category == "skills" & metric == "avg_lines" ~ "Avg Skill Size (lines)",
        category == "total" & metric == "tokens_estimate" ~ "Total Config Tokens",
        TRUE ~ paste(category, metric)
      )
    ) |>
    select(label, value)

  quality_metrics |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE)
}
```

## Evolution Over Time

```{r evolution-chart}
#| fig-width: 10
#| fig-height: 5
#| fig-alt: "Line chart showing config evolution over time with separate lines for rules, skills, hooks"

if (!is.null(config_data) && n_distinct(config_data$snapshot_date) > 1) {
  config_data |>
    filter(
      item == "all",
      metric == "count",
      category %in% c("rules", "skills", "hooks")
    ) |>
    ggplot(aes(x = snapshot_date, y = value, color = category)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Config Growth Over Time",
      x = NULL,
      y = "Count",
      color = "Category"
    ) +
    theme_dashboard
} else {
  cat("*Need multiple snapshots to show evolution. Run `config_pulse.sh` daily.*")
}
```

---

## Data Collection

Run the config pulse script daily to collect snapshots:

```bash
~/.claude/scripts/config_pulse.sh
```

Data is stored in `~/.claude/logs/config/` as Parquet files.

*Last updated: `r Sys.time()`*
