---
title: "Hover-Popup Standard: Demo"
image: man/figures/vignette-hover-popup-demo.svg
author: "John Gavin"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
vignette: >
  %\VignetteIndexEntry{Hover-Popup Standard: Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

{{< include /_includes/hover-popups.qmd >}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  message   = FALSE,
  warning   = FALSE,
  results   = "asis"
)
source(here::here("R", "hover_popup_helper.R"))
```

This vignette demonstrates the hover-popup standard introduced in
[Issue #246](https://github.com/JohnGavin/llm/issues/246). Hover (or tap on
mobile) over any underlined term to see a styled popup with full context and
an external reference link.

## Three Migrated Acronym Instances

The following paragraph migrates three acronyms from the legacy
`<abbr title="…">` pattern to the Tippy.js `tt()` pattern.

```{r demo-inline, results="asis"}
cat(
  "The ",
  tt(
    "LLM",
    paste0(
      "<b>Large Language Model</b>. A neural network trained on large text ",
      "corpora to generate, summarise, classify, and translate language. ",
      "See <a href='https://en.wikipedia.org/wiki/Large_language_model'>",
      "Wikipedia: Large Language Model</a> and the ",
      "<a href='https://anthropic.com'>Anthropic</a> model family."
    )
  ),
  " project tracks token usage and cost across ",
  tt(
    "CI",
    paste0(
      "<b>Continuous Integration</b>. An automated process that builds, tests, ",
      "and validates code on every commit. This project uses GitHub Actions for ",
      "CI. See <a href='https://docs.github.com/en/actions'>",
      "GitHub Actions documentation</a> and the ",
      "<a href='https://johngavin.github.io/llm/'>llm pkgdown site</a>."
    )
  ),
  " workflows. Pipeline outputs are validated by a ",
  tt(
    "QA",
    paste0(
      "<b>Quality Assurance</b>. Systematic checks applied to pipeline outputs ",
      "before deployment. In this project QA gates scan rendered HTML for error ",
      "patterns, missing methodology blocks, and hover-popup standard compliance. ",
      "See <a href='https://r-pkgs.org/testing-basics.html'>",
      "R Packages: Testing Basics</a> for background on automated QA in R."
    )
  ),
  " gate before each deploy.",
  sep = ""
)
```

## Usage Pattern

Include the shared partial at the top of every vignette that uses `tt()`:

````markdown
{{< include /_includes/hover-popups.qmd >}}
````

In a knitr chunk with `results = "asis"`, source the helper and call `tt()`:

```r
source(here::here("R", "hover_popup_helper.R"))
cat(tt(
  "LLM",
  paste0(
    "<b>Large Language Model</b>. ... ",
    "See <a href='https://example.com'>reference</a>."
  )
))
```

Or use Quarto's span syntax directly in markdown:

```markdown
[LLM]{.tt data-tippy-content="<b>Large Language Model</b>. ...
See <a href='https://example.com'>reference</a>."}
```

## Methodology

### What this vignette computes

This vignette does not read from the targets pipeline. It demonstrates the
`tt()` helper from `R/hover_popup_helper.R` and the `_includes/hover-popups.qmd`
shared Tippy.js partial. It serves as the pilot for the hover-popup standard
introduced in Issue #246.

### Data sources

- `R/hover_popup_helper.R`: the `tt()` function sourced directly at render time.
- `_includes/hover-popups.qmd`: the shared Tippy.js CSS + JS partial, included
  via Quarto's `{{< include >}}` directive.

### AI disclosure

This vignette was developed with assistance from Anthropic's Claude (model: Opus 4.7
and Sonnet 4.6). AI helped with code structure, prose drafting, and visualization
choices. All analytical decisions and data interpretations are the author's
responsibility.

```{r qr-footer, echo=FALSE, results="asis"}
# Minimal footer fallback when qr-footer partial is not included
invisible(NULL)
```
