---
title: "Knowledge Base Evolution Dashboard"
format:
  html:
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
---

{{< include ../_includes/qr-footer.qmd >}}


```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(scales)
library(gt)

theme_dashboard <- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold")
  )

# Try to load from local pulse data
kb_dir <- path.expand("~/.claude/logs/knowledge")
parquet_files <- list.files(kb_dir, pattern = "\\.parquet$", full.names = TRUE)

kb_data <- NULL
if (length(parquet_files) > 0 && requireNamespace("DBI", quietly = TRUE) && requireNamespace("duckdb", quietly = TRUE)) {
  tryCatch({
    con <- DBI::dbConnect(duckdb::duckdb())
    kb_data <- DBI::dbGetQuery(con, sprintf(
      "SELECT * FROM read_parquet('%s/*.parquet')",
      kb_dir
    )) |>
      mutate(
        snapshot_date = as.Date(snapshot_date),
        value = as.numeric(value)
      )
    DBI::dbDisconnect(con, shutdown = TRUE)
  }, error = function(e) NULL)
}

# Fallback: count files directly.
# In CI, ~/docs_gh/llm/knowledge resolves to a non-existent home — the data
# we want lives at <repo>/knowledge/. Probe the repo-relative path first so
# this works both in CI and locally (per #144 + portable-paths rule).
.kb_candidates <- c(here::here("knowledge"), path.expand("~/docs_gh/llm/knowledge"))
kb_path <- Find(dir.exists, .kb_candidates)
if (is.null(kb_path)) kb_path <- .kb_candidates[1]  # last-resort
wiki_dir <- file.path(kb_path, "wiki")
raw_dir <- file.path(kb_path, "raw")

if (is.null(kb_data)) {
  # Wiki pages
  if (dir.exists(wiki_dir)) {
    wiki_files <- list.files(wiki_dir, pattern = "\\.md$", full.names = TRUE)
    n_pages <- length(wiki_files)

    # Count words per page
    page_words <- sapply(wiki_files, function(f) {
      content <- paste(readLines(f, warn = FALSE), collapse = " ")
      length(strsplit(content, "\\s+")[[1]])
    })

    # Check for sources and AI-inferred
    with_sources <- sum(sapply(wiki_files, function(f) {
      any(grepl("^## Sources", readLines(f, warn = FALSE)))
    }))
    ai_inferred <- sum(sapply(wiki_files, function(f) {
      sum(grepl("AI-inferred", readLines(f, warn = FALSE)))
    }))
  } else {
    n_pages <- 0
    page_words <- numeric(0)
    with_sources <- 0
    ai_inferred <- 0
  }

  # Raw sources
  if (dir.exists(raw_dir)) {
    raw_files <- list.files(raw_dir, recursive = TRUE)
    n_sources <- length(raw_files)
  } else {
    n_sources <- 0
  }

  fallback_kb <- list(
    n_pages = n_pages,
    n_sources = n_sources,
    provenance_pct = if (n_pages > 0) round(with_sources * 100 / n_pages) else 0,
    ai_inferred = ai_inferred,
    page_words = page_words,
    with_sources = with_sources
  )
}

# Create latest dataset if kb_data exists (used by all metric chunks)
latest <- NULL
if (!is.null(kb_data)) {
  latest <- kb_data |> filter(snapshot_date == max(snapshot_date))
}
```

Track the evolution of the knowledge base: wiki pages, raw sources, confidence levels, and topic connections.

## Current Snapshot

::: {.grid}

::: {.g-col-3}
### Wiki Pages
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "page_count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_kb$n_pages)
}
```
:::

::: {.g-col-3}
### Raw Sources
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "raw", item == "all", metric == "source_count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_kb$n_sources)
}
```
:::

::: {.g-col-3}
### Provenance %
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "provenance_pct") |> pull(value)
  cat("####", n, "%")
} else {
  cat("####", fallback_kb$provenance_pct, "%")
}
```
:::

::: {.g-col-3}
### AI-Inferred
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "ai_inferred_markers") |> pull(value)
  cat("####", n, "markers")
} else {
  cat("####", fallback_kb$ai_inferred, "markers")
}
```
:::

:::

::: {.panel-tabset}

## Confidence

```{r confidence-chart}
#| fig-width: 8
#| fig-height: 4
#| fig-alt: "Stacked bar showing confidence distribution"

if (!is.null(latest)) {
  confidence <- latest |>
    filter(category == "quality", item == "all") |>
    filter(metric %in% c("fully_sourced", "partial", "unsourced")) |>
    mutate(
      label = case_when(
        metric == "fully_sourced" ~ "Fully Sourced",
        metric == "partial" ~ "Partial (has AI-inferred)",
        metric == "unsourced" ~ "Unsourced"
      ),
      label = factor(label, levels = c("Fully Sourced", "Partial (has AI-inferred)", "Unsourced"))
    )

  ggplot(confidence, aes(x = "Pages", y = value, fill = label)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 5, color = "white") +
    scale_fill_manual(values = c(
      "Fully Sourced" = "#69d4a0",
      "Partial (has AI-inferred)" = "#ffd93d",
      "Unsourced" = "#f08080"
    )) +
    coord_flip() +
    labs(
      title = "Knowledge Base Quality",
      subtitle = "Pages by confidence level",
      x = NULL,
      y = "Number of Pages",
      fill = "Confidence"
    ) +
    theme_dashboard +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
} else {
  # Fallback: categorize pages
  n_with_src <- fallback_kb$with_sources
  n_ai <- if (fallback_kb$ai_inferred > 0) min(fallback_kb$ai_inferred, fallback_kb$n_pages - n_with_src) else 0
  n_unsourced <- fallback_kb$n_pages - n_with_src

  tibble(
    label = factor(c("With Sources", "Unsourced"),
                   levels = c("With Sources", "Unsourced")),
    value = c(n_with_src, n_unsourced)
  ) |>
    ggplot(aes(x = "Pages", y = value, fill = label)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 5, color = "white") +
    scale_fill_manual(values = c("With Sources" = "#69d4a0", "Unsourced" = "#f08080")) +
    coord_flip() +
    labs(title = "Knowledge Base Quality", x = NULL, y = "Pages", fill = "Status") +
    theme_dashboard +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
}
```

## Wiki Pages

```{r wiki-pages-table}
if (!is.null(latest)) {
  wiki_pages <- latest |>
    filter(category == "wiki", metric == "words", item != "all") |>
    arrange(desc(value))

  if (nrow(wiki_pages) > 0) {
    wiki_pages |>
      gt() |>
      cols_label(item = "Page", value = "Words") |>
      cols_hide(c(snapshot_date, category, metric)) |>
      data_color(columns = value, palette = "Blues") |>
      tab_header(title = "Wiki Pages", subtitle = "Sorted by word count")
  } else {
    cat("*No wiki pages found*")
  }
} else {
  if (length(fallback_kb$page_words) > 0) {
    tibble(
      page = basename(names(fallback_kb$page_words)) |> tools::file_path_sans_ext(),
      words = as.integer(fallback_kb$page_words)
    ) |>
      arrange(desc(words)) |>
      gt() |>
      cols_label(page = "Page", words = "Words") |>
      data_color(columns = words, palette = "Blues") |>
      tab_header(title = "Wiki Pages", subtitle = "Sorted by word count")
  } else {
    cat("*No wiki pages found*")
  }
}
```

## Raw Sources

```{r raw-sources}
if (!is.null(latest)) {
  raw_summary <- latest |>
    filter(category == "raw") |>
    filter(metric == "count" | (item == "all" & metric %in% c("source_count", "bytes", "added_7d"))) |>
    mutate(
      label = case_when(
        metric == "source_count" ~ "Total Sources",
        metric == "bytes" ~ "Total Size (bytes)",
        metric == "added_7d" ~ "Added Last 7 Days",
        metric == "count" ~ paste0(".", item, " files")
      )
    ) |>
    select(label, value)

  raw_summary |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
    tab_header(title = "Raw Sources")
} else {
  tibble(
    label = c("Total Sources"),
    value = c(fallback_kb$n_sources)
  ) |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
    tab_header(title = "Raw Sources")
}
```

## Evolution

```{r evolution-chart}
#| fig-width: 10
#| fig-height: 5
#| fig-alt: "Line chart showing knowledge base growth"

if (!is.null(latest) && n_distinct(kb_data$snapshot_date) > 1) {
  kb_data |>
    filter(
      item == "all",
      metric %in% c("page_count", "source_count", "word_count")
    ) |>
    mutate(
      label = case_when(
        metric == "page_count" ~ "Wiki Pages",
        metric == "source_count" ~ "Raw Sources",
        metric == "word_count" ~ "Total Words (÷100)"
      ),
      value = if_else(metric == "word_count", value / 100, value)
    ) |>
    ggplot(aes(x = snapshot_date, y = value, color = label)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "plasma", end = 0.85) +
    labs(title = "Knowledge Base Growth", x = NULL, y = "Count", color = "Metric") +
    theme_dashboard
} else {
  cat("*Run `~/.claude/scripts/knowledge_pulse.sh` daily to collect time-series data.*

Current: ", fallback_kb$n_pages, " wiki pages, ", fallback_kb$n_sources, " raw sources")
}
```

:::

---

**Data Collection:** Run `~/.claude/scripts/knowledge_pulse.sh` daily for snapshots.

**Privacy:** Knowledge base is local-only. This dashboard shows metrics, not content.

*Last updated: `r Sys.time()`*

## Methodology

### What this vignette computes

This vignette tracks the health and growth of the knowledge base at `~/docs_gh/llm/knowledge/` — measuring wiki page count, raw source count, provenance coverage (fraction of pages with `## Sources` sections), and AI-inferred claim density. It reads time-series snapshots from `~/.claude/logs/knowledge/*.parquet` via DuckDB when available, falling back to live `list.files()` counts from `knowledge/wiki/` and `knowledge/raw/` when no snapshots exist.

### Data sources

- `~/.claude/logs/knowledge/*.parquet`: daily snapshots produced by `~/.claude/scripts/knowledge_pulse.sh`; each row is one metric for one category (wiki/raw/quality) at one snapshot date
- Live fallback: `list.files()` on `knowledge/wiki/` and `knowledge/raw/` directories — used in CI; provenance percentage computed by scanning each wiki file for `^## Sources` and `AI-inferred` patterns

### AI disclosure

This vignette was developed with assistance from Anthropic's Claude (model: Opus 4.7 and Sonnet 4.6). AI helped with code structure, prose drafting, and visualization choices. All analytical decisions and data interpretations are the author's responsibility.

{{< include ../_includes/qr-footer.qmd >}}
