---
title: "Knowledge Base Evolution Dashboard"
format:
  html:
    page-layout: full
    toc: true
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(scales)
library(gt)
library(DBI)
library(duckdb)

theme_dashboard <- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold")
  )

# Load knowledge pulse data
kb_dir <- path.expand("~/.claude/logs/knowledge")
parquet_files <- list.files(kb_dir, pattern = "\\.parquet$", full.names = TRUE)

if (length(parquet_files) > 0) {
  con <- dbConnect(duckdb())
  on.exit(dbDisconnect(con, shutdown = TRUE))

  kb_data <- dbGetQuery(con, sprintf(
    "SELECT * FROM read_parquet('%s/*.parquet')",
    kb_dir
  )) |>
    mutate(
      snapshot_date = as.Date(snapshot_date),
      value = as.numeric(value)
    )
} else {
  kb_data <- NULL
}

# Load edge list for network visualization
today <- Sys.Date()
edge_file <- file.path(kb_dir, paste0(today, "_edges.csv"))
if (file.exists(edge_file)) {
  edges <- read.csv(edge_file, stringsAsFactors = FALSE)
} else {
  edges <- NULL
}
```

Track the evolution of the knowledge base: wiki pages, raw sources, confidence levels, and topic connections.

## Current Snapshot

::: {.grid}

::: {.g-col-3}
### Wiki Pages
```{r}
#| results: asis
if (!is.null(kb_data)) {
  latest <- kb_data |> filter(snapshot_date == max(snapshot_date))
  n_pages <- latest |> filter(category == "wiki", item == "all", metric == "page_count") |> pull(value)
  cat("####", n_pages)
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-3}
### Raw Sources
```{r}
#| results: asis
if (!is.null(kb_data)) {
  n_sources <- latest |> filter(category == "raw", item == "all", metric == "source_count") |> pull(value)
  cat("####", n_sources)
} else {
  cat("#### No data")
}
```
:::
::: {.g-col-3}
### Provenance %
```{r}
#| results: asis
if (!is.null(kb_data)) {
  prov_pct <- latest |> filter(category == "wiki", item == "all", metric == "provenance_pct") |> pull(value)
  cat("####", prov_pct, "%")
} else {
  cat("#### No data")
}
```
:::

::: {.g-col-3}
### AI-Inferred
```{r}
#| results: asis
if (!is.null(kb_data)) {
  n_ai <- latest |> filter(category == "wiki", item == "all", metric == "ai_inferred_markers") |> pull(value)
  cat("####", n_ai, "markers")
} else {
  cat("#### No data")
}
```
:::

:::

## Confidence Distribution

```{r confidence-chart}
#| fig-width: 8
#| fig-height: 4
#| fig-alt: "Stacked bar showing confidence distribution: fully sourced, partial, unsourced"

if (!is.null(kb_data)) {
  confidence <- latest |>
    filter(category == "quality", item == "all") |>
    filter(metric %in% c("fully_sourced", "partial", "unsourced")) |>
    mutate(
      label = case_when(
        metric == "fully_sourced" ~ "Fully Sourced",
        metric == "partial" ~ "Partial (has AI-inferred)",
        metric == "unsourced" ~ "Unsourced"
      ),
      label = factor(label, levels = c("Fully Sourced", "Partial (has AI-inferred)", "Unsourced"))
    )

  ggplot(confidence, aes(x = "Pages", y = value, fill = label)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 5, color = "white") +
    scale_fill_manual(values = c(
      "Fully Sourced" = "#69d4a0",
      "Partial (has AI-inferred)" = "#ffd93d",
      "Unsourced" = "#f08080"
    )) +
    coord_flip() +
    labs(
      title = "Knowledge Base Quality",
      subtitle = "Pages by confidence level",
      x = NULL,
      y = "Number of Pages",
      fill = "Confidence"
    ) +
    theme_dashboard +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank()
    )
}
```

## Wiki Pages by Size

```{r wiki-pages-table}
if (!is.null(kb_data)) {
  wiki_pages <- latest |>
    filter(category == "wiki", metric == "words", item != "all") |>
    arrange(desc(value))

  if (nrow(wiki_pages) > 0) {
    wiki_pages |>
      gt() |>
      cols_label(
        item = "Page",
        value = "Words"
      ) |>
      cols_hide(c(snapshot_date, category, metric)) |>
      data_color(
        columns = value,
        palette = "Blues"
      ) |>
      tab_header(
        title = "Wiki Pages",
        subtitle = "Sorted by word count"
      )
  } else {
    cat("*No wiki pages found*")
  }
}
```

## Raw Sources Summary

```{r raw-sources}
if (!is.null(kb_data)) {
  raw_summary <- latest |>
    filter(category == "raw") |>
    filter(metric == "count" | (item == "all" & metric %in% c("source_count", "bytes", "added_7d"))) |>
    mutate(
      label = case_when(
        metric == "source_count" ~ "Total Sources",
        metric == "bytes" ~ "Total Size (bytes)",
        metric == "added_7d" ~ "Added Last 7 Days",
        metric == "count" ~ paste0(".", item, " files")
      )
    ) |>
    select(label, value)

  raw_summary |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
    tab_header(title = "Raw Sources")
}
```

## Topic Network

```{r topic-network}
#| fig-width: 8
#| fig-height: 6
#| fig-alt: "Network graph showing wiki page connections via wiki-links"

if (!is.null(edges) && nrow(edges) > 1) {
  # Simple network visualization using igraph if available
  if (requireNamespace("igraph", quietly = TRUE) &&
      requireNamespace("ggraph", quietly = TRUE)) {

    g <- igraph::graph_from_data_frame(edges, directed = TRUE)

    ggraph::ggraph(g, layout = "fr") +
      ggraph::geom_edge_link(alpha = 0.4, arrow = arrow(length = unit(3, "mm"))) +
      ggraph::geom_node_point(size = 8, color = "#4ea8de") +
      ggraph::geom_node_text(aes(label = name), repel = TRUE, size = 4) +
      labs(title = "Wiki Topic Network") +
      theme_void() +
      theme(plot.title = element_text(face = "bold", hjust = 0.5))
  } else {
    cat("*Install `igraph` and `ggraph` packages for network visualization.*")
  }
} else {
  cat("*No wiki-links found. Add `[[topic]]` links between wiki pages to build the network.*")
}
```

## Evolution Over Time

```{r evolution-chart}
#| fig-width: 10
#| fig-height: 5
#| fig-alt: "Line chart showing knowledge base growth over time"

if (!is.null(kb_data) && n_distinct(kb_data$snapshot_date) > 1) {
  kb_data |>
    filter(
      item == "all",
      metric %in% c("page_count", "source_count", "word_count")
    ) |>
    mutate(
      label = case_when(
        metric == "page_count" ~ "Wiki Pages",
        metric == "source_count" ~ "Raw Sources",
        metric == "word_count" ~ "Total Words (÷100)"
      ),
      # Scale word count for visibility
      value = if_else(metric == "word_count", value / 100, value)
    ) |>
    ggplot(aes(x = snapshot_date, y = value, color = label)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "plasma", end = 0.85) +
    labs(
      title = "Knowledge Base Growth",
      x = NULL,
      y = "Count",
      color = "Metric"
    ) +
    theme_dashboard
} else {
  cat("*Need multiple snapshots to show evolution. Run `knowledge_pulse.sh` daily.*")
}
```

---

## Data Collection

Run the knowledge pulse script daily to collect snapshots:
```bash
~/.claude/scripts/knowledge_pulse.sh
```

Data is stored in `~/.claude/logs/knowledge/` as Parquet files.

**Privacy Note:** The knowledge base is local-only. This dashboard shows metrics only, not content.

*Last updated: `r Sys.time()`*
