---
title: "Knowledge Base Evolution Dashboard"
format:
  html:
    page-layout: full
    toc: false
execute:
  echo: false
  warning: false
---

```{r setup}
#| include: false
library(dplyr)
library(ggplot2)
library(scales)
library(gt)

theme_dashboard <- theme_minimal(base_size = 14) +
  theme(
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.background = element_rect(fill = "transparent", color = NA),
    legend.background = element_rect(fill = "transparent", color = NA),
    plot.title = element_text(face = "bold")
  )

# Priority 0: CI-safe path — read pre-computed aggregates from committed RDS.
# knowledge/ is gitignored (wiki-storage-policy), so live file counts return 0
# in CI. The committed RDS ships aggregate-only stats from the last local build.
# Note: return() is invalid at chunk top-level; use Find() to select the first
# existing path, then readRDS() it outside the for loop.
kb_stats <- tryCatch({
  rds_candidates <- c(
    here::here("inst/extdata/vignettes/vig_kb_stats.rds"),
    "../inst/extdata/vignettes/vig_kb_stats.rds"
  )
  rds_hit <- Find(file.exists, rds_candidates)
  if (!is.null(rds_hit)) readRDS(rds_hit) else NULL
}, error = function(e) NULL)

# Try to load from local pulse data
kb_dir <- path.expand("~/.claude/logs/knowledge")
parquet_files <- list.files(kb_dir, pattern = "\\.parquet$", full.names = TRUE)

kb_data <- NULL
if (length(parquet_files) > 0 && requireNamespace("DBI", quietly = TRUE) && requireNamespace("duckdb", quietly = TRUE)) {
  tryCatch({
    con <- DBI::dbConnect(duckdb::duckdb())
    kb_data <- DBI::dbGetQuery(con, sprintf(
      "SELECT * FROM read_parquet('%s/*.parquet')",
      kb_dir
    )) |>
      mutate(
        snapshot_date = as.Date(snapshot_date),
        value = as.numeric(value)
      )
    DBI::dbDisconnect(con, shutdown = TRUE)
  }, error = function(e) NULL)
}

# Fallback: count files directly.
# In CI, ~/docs_gh/llm/knowledge resolves to a non-existent home — the data
# we want lives at <repo>/knowledge/. Probe the repo-relative path first so
# this works both in CI and locally (per #144 + portable-paths rule).
.kb_candidates <- c(here::here("knowledge"), path.expand("~/docs_gh/llm/knowledge"))
kb_path <- Find(dir.exists, .kb_candidates)
if (is.null(kb_path)) kb_path <- .kb_candidates[1]  # last-resort
wiki_dir <- file.path(kb_path, "wiki")
raw_dir <- file.path(kb_path, "raw")

if (is.null(kb_data)) {
  if (!is.null(kb_stats)) {
    # Priority 1: Use pre-computed RDS aggregates (CI-safe, knowledge/ gitignored).
    # page_words keys are anonymized (page_001, page_002, …) to avoid path leakage.
    fallback_kb <- list(
      n_pages       = kb_stats$n_pages,
      n_sources     = kb_stats$n_sources,
      provenance_pct = kb_stats$provenance_pct,
      ai_inferred   = kb_stats$ai_inferred_markers,
      page_words    = kb_stats$page_words,
      with_sources  = kb_stats$with_sources
    )
  } else {
    # Priority 2: Live file counts — only available when knowledge/ exists locally.
    # Wiki pages
    if (dir.exists(wiki_dir)) {
      wiki_files <- list.files(wiki_dir, pattern = "\\.md$", full.names = TRUE)
      n_pages <- length(wiki_files)

      # Count words per page
      page_words <- sapply(wiki_files, function(f) {
        content <- paste(readLines(f, warn = FALSE), collapse = " ")
        length(strsplit(content, "\\s+")[[1]])
      })

      # Check for sources and AI-inferred
      with_sources <- sum(sapply(wiki_files, function(f) {
        any(grepl("^## Sources", readLines(f, warn = FALSE)))
      }))
      ai_inferred <- sum(sapply(wiki_files, function(f) {
        sum(grepl("AI-inferred", readLines(f, warn = FALSE)))
      }))
    } else {
      n_pages <- 0
      page_words <- numeric(0)
      with_sources <- 0
      ai_inferred <- 0
    }

    # Raw sources
    if (dir.exists(raw_dir)) {
      raw_files <- list.files(raw_dir, recursive = TRUE)
      n_sources <- length(raw_files)
    } else {
      n_sources <- 0
    }

    fallback_kb <- list(
      n_pages = n_pages,
      n_sources = n_sources,
      provenance_pct = if (n_pages > 0) round(with_sources * 100 / n_pages) else 0,
      ai_inferred = ai_inferred,
      page_words = page_words,
      with_sources = with_sources
    )
  }
}

# Create latest dataset if kb_data exists (used by all metric chunks)
latest <- NULL
if (!is.null(kb_data)) {
  latest <- kb_data |> filter(snapshot_date == max(snapshot_date))
}
```

Track the evolution of the knowledge base: wiki pages, raw sources, confidence levels, and topic connections.

## Current Snapshot

::: {.grid}

::: {.g-col-3}
### Wiki Pages
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "page_count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_kb$n_pages)
}
```
:::

::: {.g-col-3}
### Raw Sources
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "raw", item == "all", metric == "source_count") |> pull(value)
  cat("####", n)
} else {
  cat("####", fallback_kb$n_sources)
}
```
:::

::: {.g-col-3}
### Provenance %
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "provenance_pct") |> pull(value)
  cat("####", n, "%")
} else {
  cat("####", fallback_kb$provenance_pct, "%")
}
```
:::

::: {.g-col-3}
### AI-Inferred
```{r}
#| results: asis
if (!is.null(latest)) {
  n <- latest |> filter(category == "wiki", item == "all", metric == "ai_inferred_markers") |> pull(value)
  cat("####", n, "markers")
} else {
  cat("####", fallback_kb$ai_inferred, "markers")
}
```
:::

:::

::: {.panel-tabset}

## Confidence

```{r confidence-chart}
#| fig-width: 8
#| fig-height: 4
#| fig-alt: "Stacked bar showing confidence distribution"

if (!is.null(latest)) {
  confidence <- latest |>
    filter(category == "quality", item == "all") |>
    filter(metric %in% c("fully_sourced", "partial", "unsourced")) |>
    mutate(
      label = case_when(
        metric == "fully_sourced" ~ "Fully Sourced",
        metric == "partial" ~ "Partial (has AI-inferred)",
        metric == "unsourced" ~ "Unsourced"
      ),
      label = factor(label, levels = c("Fully Sourced", "Partial (has AI-inferred)", "Unsourced"))
    )

  ggplot(confidence, aes(x = "Pages", y = value, fill = label)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 5, color = "white") +
    scale_fill_manual(values = c(
      "Fully Sourced" = "#69d4a0",
      "Partial (has AI-inferred)" = "#ffd93d",
      "Unsourced" = "#f08080"
    )) +
    coord_flip() +
    labs(
      title = "Knowledge Base Quality",
      subtitle = "Pages by confidence level",
      x = NULL,
      y = "Number of Pages",
      fill = "Confidence"
    ) +
    theme_dashboard +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
} else {
  # Fallback: categorize pages
  n_with_src <- fallback_kb$with_sources
  n_ai <- if (fallback_kb$ai_inferred > 0) min(fallback_kb$ai_inferred, fallback_kb$n_pages - n_with_src) else 0
  n_unsourced <- fallback_kb$n_pages - n_with_src

  tibble(
    label = factor(c("With Sources", "Unsourced"),
                   levels = c("With Sources", "Unsourced")),
    value = c(n_with_src, n_unsourced)
  ) |>
    ggplot(aes(x = "Pages", y = value, fill = label)) +
    geom_col(width = 0.6) +
    geom_text(aes(label = value), position = position_stack(vjust = 0.5), size = 5, color = "white") +
    scale_fill_manual(values = c("With Sources" = "#69d4a0", "Unsourced" = "#f08080")) +
    coord_flip() +
    labs(title = "Knowledge Base Quality", x = NULL, y = "Pages", fill = "Status") +
    theme_dashboard +
    theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
}
```

## Wiki Pages

```{r wiki-pages-table}
if (!is.null(latest)) {
  wiki_pages <- latest |>
    filter(category == "wiki", metric == "words", item != "all") |>
    arrange(desc(value))

  if (nrow(wiki_pages) > 0) {
    wiki_pages |>
      gt() |>
      cols_label(item = "Page", value = "Words") |>
      cols_hide(c(snapshot_date, category, metric)) |>
      data_color(columns = value, palette = "Blues") |>
      tab_header(title = "Wiki Pages", subtitle = "Sorted by word count")
  } else {
    cat("*No wiki pages found*")
  }
} else {
  if (length(fallback_kb$page_words) > 0) {
    tibble(
      page = basename(names(fallback_kb$page_words)) |> tools::file_path_sans_ext(),
      words = as.integer(fallback_kb$page_words)
    ) |>
      arrange(desc(words)) |>
      gt() |>
      cols_label(page = "Page", words = "Words") |>
      data_color(columns = words, palette = "Blues") |>
      tab_header(title = "Wiki Pages", subtitle = "Sorted by word count")
  } else {
    cat("*No wiki pages found*")
  }
}
```

## Raw Sources

```{r raw-sources}
if (!is.null(latest)) {
  raw_summary <- latest |>
    filter(category == "raw") |>
    filter(metric == "count" | (item == "all" & metric %in% c("source_count", "bytes", "added_7d"))) |>
    mutate(
      label = case_when(
        metric == "source_count" ~ "Total Sources",
        metric == "bytes" ~ "Total Size (bytes)",
        metric == "added_7d" ~ "Added Last 7 Days",
        metric == "count" ~ paste0(".", item, " files")
      )
    ) |>
    select(label, value)

  raw_summary |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
    tab_header(title = "Raw Sources")
} else {
  tibble(
    label = c("Total Sources"),
    value = c(fallback_kb$n_sources)
  ) |>
    gt() |>
    cols_label(label = "Metric", value = "Value") |>
    fmt_number(columns = value, decimals = 0, use_seps = TRUE) |>
    tab_header(title = "Raw Sources")
}
```

## Evolution

```{r evolution-chart}
#| fig-width: 10
#| fig-height: 5
#| fig-alt: "Line chart showing knowledge base growth"

if (!is.null(latest) && n_distinct(kb_data$snapshot_date) > 1) {
  kb_data |>
    filter(
      item == "all",
      metric %in% c("page_count", "source_count", "word_count")
    ) |>
    mutate(
      label = case_when(
        metric == "page_count" ~ "Wiki Pages",
        metric == "source_count" ~ "Raw Sources",
        metric == "word_count" ~ "Total Words (÷100)"
      ),
      value = if_else(metric == "word_count", value / 100, value)
    ) |>
    ggplot(aes(x = snapshot_date, y = value, color = label)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "plasma", end = 0.85) +
    labs(title = "Knowledge Base Growth", x = NULL, y = "Count", color = "Metric") +
    theme_dashboard
} else {
  cat("*Run `~/.claude/scripts/knowledge_pulse.sh` daily to collect time-series data.*

Current: ", fallback_kb$n_pages, " wiki pages, ", fallback_kb$n_sources, " raw sources")
}
```

:::

---

**Data Collection:** Run `~/.claude/scripts/knowledge_pulse.sh` daily for snapshots.

**Privacy:** Knowledge base is local-only. This dashboard shows metrics, not content.

*Last updated: `r Sys.time()`*

## Architecture & Intent

The knowledge base (KB) is a **private "second brain"** — a local-only git repository at `~/docs_gh/llm/knowledge/` that accumulates raw source material, distilled wiki pages, and outputs, all under a `PRIVATE` marker that blocks any accidental push to GitHub (per the [`knowledge-base-wiki`](https://github.com/JohnGavin/llm/blob/main/.claude/skills/knowledge-base-wiki.md) skill and the [`wiki-storage-policy`](https://github.com/JohnGavin/llm/blob/main/.claude/rules/wiki-conventions.md#L17) rule). The dashboard on this page and the daily [Knowledge Base (KB) digest email](https://github.com/JohnGavin/llm/issues/298) summarise *how the KB grows* — tracking counts, provenance coverage, and topic connections — without ever exposing its contents. This is the central design constraint: CI pipelines, dashboards, and emails see aggregates and titles only; raw bodies never cross the local boundary.

### Knowledge Base Data Flow

The diagram below shows the three-layer structure and where the privacy boundary sits. Clickable nodes link to the relevant source files and rules.

```{mermaid}
%%| fig-alt: "Knowledge base data flow: raw (append-only) feeds wiki (provenance + cross-links) feeds outputs. A privacy boundary prevents raw content from reaching CI, the dashboard, or the digest email. knowledge_pulse.sh extracts aggregate metrics only."

flowchart LR
  subgraph LOCAL ["🔒 Local only (PRIVATE + pre-push block)"]
    RAW["raw/\n(append-only)"]
    WIKI["wiki/\n(provenance + [[links]])"]
    OUT["outputs/"]
    RAW --> WIKI
    WIKI --> OUT
  end

  subgraph METRICS ["Aggregate metrics only"]
    PULSE["knowledge_pulse.sh\n(counts · titles · themes)"]
    RDS["vig_kb_stats.rds\n(anonymised aggregates)"]
    PULSE --> RDS
  end

  WIKI -->|"metrics only\n(no bodies)"| PULSE

  subgraph PUBLISH ["Public / CI"]
    DASH["KB-evolution\ndashboard"]
    EMAIL["daily digest\n(titles + counts)"]
    RDS --> DASH
    PULSE -->|"summaries only"| EMAIL
  end

  PRIVACY["🚫 Raw content\nNEVER here"]
  PUBLISH -.->|blocked| PRIVACY

  click RAW "https://github.com/JohnGavin/llm/blob/main/.claude/rules/wiki-conventions.md#L44" _blank
  click WIKI "https://github.com/JohnGavin/llm/blob/main/.claude/rules/wiki-conventions.md#L63" _blank
  click PULSE "https://github.com/JohnGavin/llm/blob/main/.claude/scripts/knowledge_pulse.sh#L1" _blank
  click RDS "https://github.com/JohnGavin/llm/issues/122" _blank
  click EMAIL "https://github.com/JohnGavin/llm/issues/298" _blank
  click DASH "https://johngavin.github.io/llm/articles/knowledge-evolution.html" _blank
```

### Key ideas for a new reader

- **Local-only git with hard block**: `knowledge/` has no remote and a `pre-push` hook that aborts if the `PRIVATE` marker is present — a push to GitHub is structurally impossible, not just discouraged.
- **`raw/` is append-only**: files in `raw/` are write-once; editing an existing file corrupts the provenance chain and creates an AI-output-as-input feedback loop. The `file_protection.sh` hook enforces this.
- **Every wiki page requires `## Sources`**: a page without a `## Sources` section is flagged by `/wiki-health` and the T1 on-write hook. Sources link back to specific lines in `raw/` files.
- **AI-inferred claims are tagged `> ⚠ AI-inferred:`**: this marks synthesised conclusions that are not direct quotes, allowing readers (and future wiki authors) to distinguish evidence strength.
- **Confidence markers can be downgraded**: a claim promoted to "source-stated" can be demoted back to "AI-inferred" if closer inspection reveals it was a synthesis. The health metric targets >70% source-stated, <20% AI-inferred.
- **The dashboard and digest ship aggregates and titles, never bodies**: `knowledge_pulse.sh` records counts, line deltas, and page-title lists — not page content. The committed `vig_kb_stats.rds` anonymises even file paths (`page_001`, `page_002`, …).

### Questions a new reader should ask

- *What is private vs publishable here?* Counts, titles, and themes are publishable. Page bodies, raw source text, and any Personally Identifiable Information (PII) within raw files are strictly local.
- *Where did this claim come from?* Every wiki claim should trace to a `## Sources` entry linking to a `raw/` file with a line anchor. If it does not, the provenance is incomplete.
- *How confident is it?* Look for `> ⚠ AI-inferred:` (synthesised) vs unmarked text (direct quote/paraphrase). The `consensus_level` frontmatter field gives the page-level view.
- *Why does CI show aggregates only?* Because `knowledge/` is gitignored — CI never sees the file contents. The dashboard reads the committed `vig_kb_stats.rds`, which ships aggregate-only stats from the last local `tar_make()` run.
- *How is the digest computed without leaking content?* [`knowledge_pulse.sh`](https://github.com/JohnGavin/llm/blob/main/.claude/scripts/knowledge_pulse.sh#L1) runs locally, extracts only metrics and titles, and writes them to `~/.claude/logs/knowledge/*.parquet`. The digest email reads those summaries — raw bodies never enter the email pipeline.

### Structure & application

Raw notes or transcripts land in `raw/` (append-only). The `wiki-curator` agent distils these into `wiki/*.md` pages, each with frontmatter, a `## Sources` section, and `[[topic]]` cross-links. `knowledge_pulse.sh` runs nightly and writes a row-per-metric snapshot to a local Parquet file. `tar_make()` reads those snapshots (or the live file counts as fallback) and writes `vig_kb_stats.rds` into `inst/extdata/vignettes/`. This vignette reads that committed RDS — so the dashboard works in CI without ever touching the local KB.

A daily digest row maps directly to a dashboard panel: the "Wiki Pages" KPI comes from `vig_kb_stats.rds$n_pages`; the "Evolution" tab time-series comes from `~/.claude/logs/knowledge/*.parquet`; the "Confidence" bar comes from the `quality` category rows in the same Parquet files. Running `/wiki-health` triggers the same provenance and staleness checks that feed both the digest and the dashboard.

**Related:** [#298 KB digest email](https://github.com/JohnGavin/llm/issues/298) · [#301 config-side sibling](https://github.com/JohnGavin/llm/issues/301) · [`knowledge-base-wiki` skill](https://github.com/JohnGavin/llm/blob/main/.claude/skills/knowledge-base-wiki.md) · [`wiki-storage-policy` rule](https://github.com/JohnGavin/llm/blob/main/.claude/rules/wiki-conventions.md#L17) · `/wiki-health` command

## Methodology

### What this vignette computes

This vignette tracks the health and growth of the knowledge base at `~/docs_gh/llm/knowledge/` — measuring wiki page count, raw source count, provenance coverage (fraction of pages with `## Sources` sections), and AI-inferred claim density. It reads time-series snapshots from `~/.claude/logs/knowledge/*.parquet` via DuckDB when available, falling back to live `list.files()` counts from `knowledge/wiki/` and `knowledge/raw/` when no snapshots exist.

### Data sources

- `inst/extdata/vignettes/vig_kb_stats.rds`: pre-computed aggregate stats from the `vig_kb_stats` targets pipeline target, committed to the repository; read first in all environments including CI (where `knowledge/` is gitignored); page word counts are anonymized (`page_001`, `page_002`, …) to avoid leaking file paths
- `~/.claude/logs/knowledge/*.parquet`: daily snapshots produced by `~/.claude/scripts/knowledge_pulse.sh`; each row is one metric for one category (wiki/raw/quality) at one snapshot date; used when available locally for richer time-series charts
- Live fallback: `list.files()` on `knowledge/wiki/` and `knowledge/raw/` directories — used only when both the RDS and parquet paths are unavailable; provenance percentage computed by scanning each wiki file for `^## Sources` and `AI-inferred` patterns

### AI disclosure

This vignette was developed with assistance from Anthropic's Claude (model: Opus 4.7 and Sonnet 4.6). AI helped with code structure, prose drafting, and visualization choices. All analytical decisions and data interpretations are the author's responsibility.

{{< include ../_includes/qr-footer.qmd >}}
