---
title: "Project Telemetry"
author: "John Gavin"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    page-layout: full
    code-fold: true
vignette: >
  %\VignetteIndexEntry{Project Telemetry}
  %\VignetteEngine: quarto::html
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  echo = FALSE
)

has_targets <- requireNamespace("targets", quietly = TRUE)

# Discover target store
tar_store <- tryCatch({
  if (file.exists(here::here("_targets"))) {
    here::here("_targets")
  } else if (file.exists("_targets")) {
    "_targets"
  } else {
    NULL
  }
}, error = function(e) NULL)

# Safe target reader - returns NULL if target unavailable
safe_tar_read <- function(name) {
  if (is.null(tar_store)) return(NULL)
  tryCatch(
    targets::tar_read_raw(name, store = tar_store),
    error = function(e) NULL
  )
}
```

# LLM Usage & Costs {#llm-usage-costs}

This section tracks costs, token usage, and limits.

```{r dashboard-status}
safe_tar_read("vig_usage_summary")
```

::: {.panel-tabset}

## Cost Trends

```{r cost-trend, fig.cap="Daily LLM costs over time with LOESS smoothing trend line."}
safe_tar_read("vig_cost_trend_plot")
```

## Cumulative Cost

```{r cumulative-cost, fig.cap="Cumulative LLM spending over time showing total expenditure trajectory."}
safe_tar_read("vig_cumulative_cost_plot")
```

## Breakdowns

```{r breakdowns-combined, fig.height=10, fig.cap="Top: Cost breakdown by model. Bottom: Token usage by type (input, output, cache)."}
safe_tar_read("vig_breakdowns_plot")
```

## Gemini Analytics

```{r gemini-analysis, fig.cap="Gemini daily costs from the Gemini usage database."}
safe_tar_read("vig_gemini_plot")
```

:::

# Session Efficiency {#session-efficiency}

Analysis of session durations and cost efficiency, including Claude Max5 blocks.

::: {.panel-tabset}

## Duration Trends

```{r duration-trend, fig.cap="Average session duration over time, colored by recency period."}
safe_tar_read("vig_duration_trend_plot")
```

## Cost Efficiency

```{r cost-eff, fig.cap="Cost per minute over time showing session cost efficiency trends."}
safe_tar_read("vig_cost_efficiency_plot")
```

## Cost vs Duration

```{r cost-duration, fig.cap="Scatter plot of cost intensity vs session duration. Longer sessions tend to be more cost efficient."}
safe_tar_read("vig_cost_duration_plot")
```

## Model Breakdown

```{r model-breakdown, fig.height=8, fig.cap="Cost efficiency faceted by Claude model variant."}
safe_tar_read("vig_model_session_plot")
```

## Max5 Block History {#max5-block-history}

Recent history of usage blocks. "Max5" refers to the **Claude 5-hour usage limit** window^[The "Max5" metric tracks token consumption within the rolling 5-hour window enforced by Anthropic for Claude models.].

<div title="Max5 refers to Claude's 5-hour usage window where token limits are strictly enforced.">
*(Hover for info: What is Max5?)*
</div>

```{r max5-table}
safe_tar_read("vig_max5_table")
```

:::

# Pipeline Metrics {#pipeline-metrics}

Targets pipeline composition: plans, target counts, heaviest targets by size and compute time.

::: {.panel-tabset}

## Plans & Targets

```{r pipeline-plans}
safe_tar_read("vig_pipeline_plans_table")
```

## Top by Size

```{r pipeline-top-size}
safe_tar_read("vig_pipeline_top_size_table")
```

## Top by Compute Time

```{r pipeline-top-time}
safe_tar_read("vig_pipeline_top_time_table")
```

:::

# GitHub Activity {#github-activity}

Repository timeline, commit velocity, issues, pull requests, and CI workflows.

::: {.panel-tabset}

## Commit Velocity

```{r commit-velocity}
safe_tar_read("vig_commit_velocity_table")
```

## Issues & PRs

```{r github-activity}
safe_tar_read("vig_github_activity_table")
```

## CI Workflow Runtimes

```{r ci-plot, fig.cap="Box plot of successful GitHub Actions workflow runtimes by workflow name."}
safe_tar_read("vig_workflow_plot")
```

## Git History

```{r git-hist, fig.cap="Daily commit frequency over the last 100 commits."}
safe_tar_read("vig_git_history_plot")
```

:::

# Project Structure {#project-structure}

Codebase metrics, file organization, and repository health.

::: {.panel-tabset}

## Codebase

```{r codebase-metrics}
safe_tar_read("vig_codebase_metrics")
```

## File Counts

```{r file-counts}
safe_tar_read("vig_file_counts_table")
```

## GitHub Stats

```{r github-stats}
safe_tar_read("vig_github_stats_table")
```

:::

# Prediction Calibration {#prediction-calibration}

Cross-project calibration metrics for Claude Code probability assessments.
Brier score: 0 = perfect, 0.25 = uninformative baseline.

::: {.panel-tabset}

## By Project

```{r pred-by-project}
safe_tar_read("vig_pred_by_project_table")
```

## Calibration Buckets

```{r pred-calibration}
safe_tar_read("vig_pred_global_calibration_table")
```

## Rolling Brier Score

```{r pred-rolling-brier, fig.cap="Cumulative Brier score across all projects over time. Lower is better; 0.25 is the uninformative baseline."}
safe_tar_read("vig_pred_global_brier_plot")
```

## Predicted vs Actual

```{r pred-success-rate, fig.cap="Comparison of mean predicted probability vs actual success rate by project."}
safe_tar_read("vig_pred_success_rate_plot")
```

:::
