---
title: "The Claude Code Config Stack"
format:
  closeread-html:
    toc: false
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
---

<style>
.cr-section.sidebar-left .cr-sidebar-col { flex: 0 0 55% !important; max-width: 55% !important; }
.cr-section.sidebar-left .cr-narrative-col { flex: 0 0 45% !important; max-width: 45% !important; }
.sticky pre, .sticky table { font-size: 0.9rem; }
.sticky table td, .sticky table th { padding: 4px 8px; }
.cr-narrative-col p { font-size: 0.95rem; line-height: 1.6; }
.sticky table td code, .sticky table td .cr-highlight { color: #e6db74; font-weight: 600; }
</style>

```{r setup}
#| include: false
safe_tar_read <- function(name) {
  store <- tryCatch({
    owd <- setwd(here::here())
    on.exit(setwd(owd))
    normalizePath(targets::tar_config_get("store"))
  }, error = function(e) NULL)
  if (!is.null(store)) {
    obj <- tryCatch(targets::tar_read_raw(name, store = store), error = function(e) NULL)
    if (!is.null(obj)) return(obj)
  }
  for (d in c("../inst/extdata/vignettes", "inst/extdata/vignettes",
              file.path(here::here(), "inst/extdata/vignettes"))) {
    rds <- file.path(d, paste0(name, ".rds"))
    if (file.exists(rds)) return(readRDS(rds))
  }
  NULL
}

counts <- safe_tar_read("vig_cr_infra_counts")
if (is.null(counts)) counts <- list(n_rules=61, n_skills=63, n_agents=12, n_hooks=7, n_commands=15, n_memory=14)
```

:::{.cr-section}

:::{#cr-entry .sticky}
```
# ~/.claude/CLAUDE.md

## Core Rules
**Session Start:** echo $IN_NIX_SHELL
**Git:** gert::git_add(), git_commit(), git_push()
**Nix:** nix-shell /absolute/path/default.nix --run "cmd"
**Errors:** NEVER speculate. READ error, QUOTE it.

## Agents (12)
## Skills (63)
## Commands (15)
## Rules (61)
## Memory (14 files)
## Hooks (7 scripts)
```
:::

Every Claude Code session begins by loading `~/.claude/CLAUDE.md`. This file is the entry point — the first thing read before any tool is called, any file touched, any code written.

It is not a README. It is a **constitution**: a concise set of constraints, pointers, and defaults that shape every decision the model makes. From the outside, Claude appears to respond to prompts. From the inside, it is navigating a graph of rules, skills, agents, and memory — all loaded from this one file.

@cr-entry

The config stack has six layers: rules that forbid, skills that enable, agents that specialize, memory that persists, commands that trigger, and hooks that automate. Each layer adds a dimension of control.

@cr-entry

:::

:::{.cr-section}

:::{#cr-tree .sticky}
```{r tree-block}
#| results: asis
cat(paste0(
'```\n',
'~/.claude/\n',
'\u251c\u2500\u2500 CLAUDE.md              \u2190 entry point\n',
'\u251c\u2500\u2500 rules/                 \u2190 ', counts$n_rules, ' constraint files\n',
'\u2502   \u251c\u2500\u2500 btw-timeouts.md\n',
'\u2502   \u251c\u2500\u2500 git-no-compound-cd.md\n',
'\u2502   \u2514\u2500\u2500 ...\n',
'\u251c\u2500\u2500 skills/                \u2190 ', counts$n_skills, ' capability modules\n',
'\u2502   \u251c\u2500\u2500 adversarial-qa/\n',
'\u2502   \u251c\u2500\u2500 shiny-bslib/\n',
'\u2502   \u2514\u2500\u2500 ...\n',
'\u251c\u2500\u2500 agents/                \u2190 ', counts$n_agents, ' specialist definitions\n',
'\u2502   \u251c\u2500\u2500 critic.md\n',
'\u2502   \u251c\u2500\u2500 quick-fix.md\n',
'\u2502   \u2514\u2500\u2500 ...\n',
'\u251c\u2500\u2500 commands/              \u2190 ', counts$n_commands, ' slash commands\n',
'\u2502   \u251c\u2500\u2500 check.md\n',
'\u2502   \u251c\u2500\u2500 session-end.md\n',
'\u2502   \u2514\u2500\u2500 ...\n',
'\u251c\u2500\u2500 hooks/                 \u2190 ', counts$n_hooks, ' event scripts\n',
'\u2502   \u251c\u2500\u2500 session_init.sh\n',
'\u2502   \u2514\u2500\u2500 ...\n',
'\u2514\u2500\u2500 projects/*/memory/     \u2190 ', counts$n_memory, ' persistent memory files\n',
'    \u251c\u2500\u2500 MEMORY.md\n',
'    \u2514\u2500\u2500 ...\n',
'```\n'
))
```
:::

The file tree shows the physical layout. Every `.md` file in `rules/` is a constraint. Every directory in `skills/` is a capability module. Every `.md` in `agents/` is a specialist definition.

@cr-tree

The tree is shallow — one level deep for each config type. This flatness is intentional. Rules don't have sub-rules. Skills don't nest inside skills. The complexity lives in how these flat files interact, not in their directory structure.

@cr-tree

:::

:::{.cr-section}

:::{#cr-rules .sticky}
```{r rules-table}
#| results: asis
rule_data <- safe_tar_read("vig_cr_rule_categories")
if (is.null(rule_data)) {
  rule_data <- tibble::tibble(
    category    = c("Core workflow", "Nix / environment", "Data quality", "Statistics",
                    "Visualization", "Quarto / vignettes", "Shiny / web", "Git / safety",
                    "Knowledge base", "Other"),
    count       = c(5L, 3L, 7L, 6L, 5L, 7L, 3L, 2L, 4L, 19L),
    example_rule = c(
      "systematic-debugging, verification-before-completion",
      "nix-agent-shell-protocol, nix-nested-shell-isolation",
      "data-validation-timeseries, duckdplyr-not-sql",
      "statistical-reporting, look-ahead-bias-prevention",
      "visualization-standards, diagram-generation",
      "quarto-vignette-format, vignette-targets-export",
      "shinylive-webr-nonblocking, module-isolation",
      "git-no-compound-cd, safe-deletion",
      "wiki-staleness-check, raw-folder-readonly",
      "btw-timeouts, confidence-markers, ..."
    )
  )
}
knitr::kable(rule_data, format = "html",
  col.names = c("Category", "Count", "Examples"),
  caption = paste0("Rule categories across ", counts$n_rules, " config files"),
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

Rules are not guidelines. They are constraints — hard stops that apply regardless of context, user request, or apparent convenience. Right now there are `r counts$n_rules` of them.

@cr-rules

Each rule file has the same anatomy: a trigger clause ("When This Applies"), a critical violation, the correct substitution pattern, and a rationale. The trigger clause matters most — it defines precisely when the rule activates so the model doesn't apply it everywhere indiscriminately.

@cr-rules

Some rules exist because of a single incident: a 522 MB worktree deleted without checking size or age (`safe-deletion`). A git operation blocked by a bare-repo guard that no permission setting could bypass (`git-no-compound-cd`). Rules written in blood are the ones that stick.

@cr-rules

:::

:::{.cr-section}

:::{#cr-rule-anatomy .sticky}
```
# Rule: No `cd &&` Compound Commands       ← 1. Title

## When This Applies                        ← 2. Trigger clause
Every time Claude calls a Bash tool that
operates on files or repos in any directory
other than the current working directory.

## CRITICAL: Never Use `cd <dir> && git …`  ← 3. Violation
Claude Code has a hardcoded safety guard
that prompts for approval whenever a Bash
call combines `cd` with `git`.

## Mandatory Substitution                   ← 4. Fix
| Wrong             | Right            |
|-------------------|------------------|
| cd ~/r && git log | git -C ~/r log   |
| cd ~/r && git add | git -C ~/r add   |
```
:::

This is what a real rule looks like from the inside. Four parts, always the same structure: [title]{.cr-highlight}, [trigger]{.cr-highlight}, [violation]{.cr-highlight}, [fix]{.cr-highlight}.

@cr-rule-anatomy

The [trigger clause]{.cr-highlight} is the sharpest part. "Every time Claude calls a Bash tool that operates on files or repos" — not "sometimes" or "when convenient." The precision matters because the model interprets ambiguity as latitude.

@cr-rule-anatomy

The [substitution table]{.cr-highlight} provides a mechanical fix. No judgment needed. `cd ~/repo && git status` becomes `git -C ~/repo status`. Every git subcommand accepts `-C`. The table makes the pattern undeniable.

@cr-rule-anatomy

:::

:::{.cr-section}

:::{#cr-skills .sticky}
```{r skills-table}
#| results: asis
skill_data <- safe_tar_read("vig_cr_skill_categories")
if (is.null(skill_data)) {
  skill_data <- tibble::tibble(
    category = c("Mandatory", "R package development", "Data & analysis",
                 "Targets & pipelines", "Shiny & web", "Quarto & docs",
                 "DevOps & CI", "Project management", "AI / LLM tools", "Prose quality"),
    count    = c(8L, 12L, 9L, 4L, 8L, 7L, 3L, 8L, 5L, 1L),
    purpose  = c(
      "Always applied: adversarial-qa, quality-gates, r-package-workflow ...",
      "cli, dplyr, vctrs, S7, lifecycle, testthat ...",
      "EDA, pointblank, DuckDB, modeling, genomics ...",
      "pipeline spec, vignettes, crew, parallel",
      "bslib, async, modules, Shinylive, plumber2 ...",
      "websites, dashboards, dynamic content, WebR ...",
      "GitHub Actions, pkgdown, static API",
      "plans, review, requirements, context control ...",
      "Gemini CLI, HuggingFace, MCP servers, hooks",
      "deslop — removes AI writing patterns"
    )
  )
}
knitr::kable(skill_data, format = "html",
  col.names = c("Category", "Count", "Purpose"),
  caption = paste0("Skill modules by domain (", counts$n_skills, " total)"),
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

If rules define what not to do, skills define what to do. There are `r counts$n_skills` skill modules, each invoked by name: `adversarial-qa` for attack-based testing, `deslop` for removing AI writing patterns, `r-package-workflow` for the nine-step PR process.

@cr-skills

Skills come in two varieties: mandatory and situational. The eight mandatory skills load for every task — quality gates, adversarial QA, TDD, the R package workflow. They are always active. Situational skills activate when the user's request matches a trigger phrase.

@cr-skills

The trigger mechanism is intentionally loose. "Run tests" activates `test-driven-development`. "Shinylive" activates `shinylive-deployment`. "Make it sound human" activates `deslop`. The model pattern-matches on intent, not exact wording.

@cr-skills

:::

:::{.cr-section}

:::{#cr-agents .sticky}
```{r agents-table}
#| results: asis
agent_data <- safe_tar_read("vig_cr_agent_tiers")
if (is.null(agent_data)) {
  agent_data <- tibble::tibble(
    agent   = c("quick-fix", "critic", "fixer", "r-debugger", "targets-runner",
                "reviewer", "nix-env", "shiny-async-debugger", "data-quality-guardian",
                "data-engineer", "shinylive-builder", "wiki-curator"),
    model   = c("haiku", rep("sonnet", 11L)),
    trigger = c(
      "Typos, renames, version bumps, URL fixes",
      "Read-only adversarial review",
      "Apply fixes from critic reports",
      '"run tests", "fix test failure"',
      '"tar_make", "pipeline failed"',
      '"review this PR/code"',
      '"nix shell broken", "package missing"',
      '"ExtendedTask", "crew bug"',
      '"validate data", "pointblank"',
      '"dbt", "SQL pipeline"',
      '"shinylive", "WASM build"',
      '"compile wiki from raw"'
    )
  )
} else {
  agent_data <- agent_data[, c("agent", "model", "description")]
}
knitr::kable(agent_data, format = "html",
  col.names = c("Agent", "Model", "Trigger"),
  caption = "Agent model tiers and delegation triggers",
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

The twelve agents exist because not every task needs the same capability — or the same cost. The `quick-fix` agent runs on [haiku]{.cr-highlight} for mechanical edits. Every other specialist runs on [sonnet]{.cr-highlight}. The orchestrator runs on [opus]{.cr-highlight}.

@cr-agents

Auto-delegation is mandatory. If the task involves a single file, fewer than five line changes, and no reasoning about correctness, it goes to haiku. If it matches a specialist trigger, it goes to that specialist. Only multi-file architecture decisions and synthesis tasks stay with the orchestrator.

@cr-agents

The critic and fixer are complementary: critic is read-only and adversarial, fixer is read-write and applies what critic finds. They run in sequence — critic first, fixer second, critic again to verify. Three rounds maximum before escalating to the user.

@cr-agents

:::

:::{.cr-section}

:::{#cr-memory .sticky}
```
# ~/.claude/projects/.../memory/MEMORY.md

## Architecture (see architecture.md)
## MCP r-btw: NEVER Call Directly
## Agent Patterns (see agent-patterns.md)
## CI Strategy (see ci-strategy.md)
## Nix Operations (see nix-operations.md)
## Shinylive (see shinylive-issues.md)
## btw MCP Tools
## pkgctx
## Tool Preferences (see tool-preferences.md)
## Session Conventions
## Safe Deletion
## NEVER Edit default.nix Directly
## Nix Shell Portability
## Never `cd && git`
## Knowledge Base Discipline
## GitHub Pages User Sites
## Config Migration (2026-03-09)
```
:::

Memory persists what conversation context cannot. The `r counts$n_memory` memory files capture lessons learned, failed approaches, and standing conventions — facts that need to survive across sessions.

@cr-memory

There are four kinds of memory here. [Architecture]{.cr-highlight} docs explain how the system is structured. [Feedback]{.cr-highlight} files record incidents: what went wrong, what the fix was, why the rule now exists. [Convention]{.cr-highlight} files capture standing decisions. The [index]{.cr-highlight} ties them together.

@cr-memory

The memory system works because it is curated, not accumulated. Each feedback file covers exactly one incident. Each convention file covers exactly one domain. The index is under twenty lines. When the index gets long, something has gone wrong.

@cr-memory

:::

:::{.cr-section}

:::{#cr-commands .sticky}
```{r commands-table}
#| results: asis
cmd_data <- safe_tar_read("vig_cr_commands")
if (is.null(cmd_data)) {
  cmd_data <- tibble::tibble(
    command  = c("/hi", "/bye", "/check", "/ctx-check", "/pr-status", "/cleanup",
                 "/issue-triage", "/new-issue", "/wiki-health", "/wiki-promote", "/write-alt-text"),
    alias    = c("/session-start", "/session-end", "\u2014", "\u2014", "\u2014", "\u2014",
                 "/triage", "\u2014", "\u2014", "\u2014", "\u2014"),
    does     = c(
      "Read CHANGELOG, check git status, open issues",
      "Commit, append CHANGELOG, push, wiki staleness check",
      "R CMD check + code sweep with ast-grep rules",
      "Check ctx.yaml coverage for all exported functions",
      "Check open PRs and CI status",
      "Review and simplify work (critic + fixer pass)",
      "List GitHub issues by difficulty",
      "Create GitHub issue with branch",
      "Validate knowledge base wiki frontmatter and links",
      "Promote an output into the wiki",
      "Generate alt text for all figures"
    )
  )
}
knitr::kable(cmd_data, format = "html",
  col.names = c("Command", "Alias", "What it does"),
  caption = "Slash commands for session management",
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

Commands are skills with a slash in front. `/hi` triggers `session-start`, which reads CHANGELOG, checks git status, and orients the session. `/bye` triggers `session-end`, which commits, appends the CHANGELOG, and checks for stale wiki pages.

@cr-commands

The `/hi` to `/bye` lifecycle structures the session. Not every session follows it — sometimes the user drops in mid-task — but the pattern creates a rhythm: start with context, end with a record. The CHANGELOG is a lab notebook, not release notes.

@cr-commands

Some commands are diagnostic: `/check` runs the full quality gate sweep, `/ctx-check` verifies API context coverage, `/pr-status` polls GitHub Actions. These are not automated — they run when called. Automation lives in hooks.

@cr-commands

:::

:::{.cr-section}

:::{#cr-hooks .sticky}
```{r hooks-table}
#| results: asis
hook_data <- safe_tar_read("vig_cr_hook_lifecycle")
if (is.null(hook_data)) {
  hook_data <- tibble::tibble(
    script = c("session_init.sh", "context_survival.sh", "file_protection.sh",
               "context_monitor.sh", "wiki_health_onwrite.sh", "session_stop.sh"),
    event  = c("SessionStart", "PreCompact + compact/resume", "PreToolUse: Edit | Write",
               "PostToolUse: Bash | Task", "PostToolUse: Edit | Write", "Stop"),
    action = c(
      "Load context, check Nix shell, orient session",
      "Preserve critical state before context window compaction",
      "Block writes to raw/ (append-only), flag protected files",
      "Track context window usage, warn at thresholds",
      "T1 health check on every wiki write (frontmatter, Sources)",
      "Commit pending changes, append CHANGELOG"
    )
  )
}
knitr::kable(hook_data, format = "html",
  col.names = c("Script", "Event", "Action"),
  caption = "Hook lifecycle — event-driven automation",
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

Hooks fire on events. Six scripts cover the session lifecycle: initialization, context survival, file protection, context monitoring, wiki health, and session stop. They are invisible when working correctly — you notice them only when they block something.

@cr-hooks

[`file_protection.sh`]{.cr-highlight} is the most important. It fires on every `Edit` or `Write` call and checks whether the target path is in `raw/`. Raw files are append-only — not deletable, not replaceable, only extendable. The hook enforces this at the filesystem level.

@cr-hooks

[`context_survival.sh`]{.cr-highlight} fires before context window compaction. It writes critical session state — current task, key decisions, failed approaches — to `CURRENT_WORK.md`. When the context is rebuilt from scratch, this file is the bridge that prevents repeated mistakes.

@cr-hooks

:::

:::{.cr-section}

:::{#cr-hook-anatomy .sticky}
```bash
#!/usr/bin/env bash
# file_protection.sh
# Hook: PreToolUse (Edit, Write)
# Exit 2 = BLOCK. Exit 0 = WARN.

INPUT=$(cat)
FILE_PATH=$(echo "$INPUT" | sed -n '...')

# Auto-generated files: BLOCK (exit 2)
BLOCK_PATTERNS=("NAMESPACE" "man/")
for pattern in "${BLOCK_PATTERNS[@]}"; do
  if [[ "$FILE_PATH" == *"$pattern"* ]]; then
    echo "BLOCKED: $FILE_PATH is auto-generated"
    exit 2  # ← hard block, tool call rejected
  fi
done

# raw/ folders: append-only
if [[ "$FILE_PATH" == *"/raw/"* ]]; then
  echo "BLOCKED: raw/ is append-only"
  exit 2
fi
```
:::

Here is `file_protection.sh` — the hook that enforces append-only on `raw/` and blocks edits to auto-generated files. It reads the tool call's [JSON payload via stdin]{.cr-highlight}, extracts the file path, and pattern-matches.

@cr-hook-anatomy

The exit codes are the contract: [exit 2]{.cr-highlight} is a hard block — the tool call is rejected and the model must find another way. Exit 0 passes through. There is no "soft warning" — the hook either blocks or it doesn't.

@cr-hook-anatomy

Two protection tiers. [NAMESPACE and man/]{.cr-highlight} are auto-generated by `devtools::document()` — editing them directly is wasted work because the next documentation rebuild overwrites everything. [`raw/`]{.cr-highlight} is the knowledge base source of truth — append-only, never edited, never deleted.

@cr-hook-anatomy

:::

:::{.cr-section}

:::{#cr-composition .sticky}
```{r summary-table}
#| results: asis
summary_data <- tibble::tibble(
  layer = c("Rules", "Skills", "Agents", "Memory", "Commands", "Hooks"),
  count = c(counts$n_rules, counts$n_skills, counts$n_agents,
            counts$n_memory, counts$n_commands, counts$n_hooks),
  role  = c(
    "Constrain every action (what not to do)",
    "Supply domain knowledge (what to do and how)",
    "Delegate to specialists (who does the work)",
    "Carry lessons from past sessions",
    "Trigger structured workflows on demand",
    "Automate enforcement on tool events"
  )
)
knitr::kable(summary_data, format = "html",
  col.names = c("Layer", "Count", "Role in a request"),
  caption = "The six layers of the config stack",
  table.attr = 'class="table table-sm" style="font-size: 0.9rem;"')
```
:::

When you type a request, all six layers activate simultaneously. [Rules]{.cr-highlight} load as background constraints. Mandatory [skills]{.cr-highlight} attach their protocols. The model identifies any matching specialist [agents]{.cr-highlight} and queues delegation decisions. [Memory]{.cr-highlight} files surface relevant past incidents. [Commands]{.cr-highlight} check for slash-trigger patterns. [Hooks]{.cr-highlight} arm themselves against the tool calls about to come.

@cr-composition

Then the work begins. A file edit trips `file_protection.sh`. A Bash call updates context usage in `context_monitor.sh`. A git operation checks the `git-no-compound-cd` rule. A quality gate scores the result against the rubric. A wiki write fires the T1 health check.

@cr-composition

None of these layers operates in isolation. A skill invokes an agent. An agent applies rules. A hook validates what the agent did. Memory records what went wrong. A command structures the session that started it all. The stack is not a hierarchy — it is a web.

@cr-composition

`r counts$n_rules` rules. `r counts$n_skills` skills. `r counts$n_agents` agents. `r counts$n_memory` memory files. `r counts$n_commands` commands. `r counts$n_hooks` hooks. This is what it takes to make a language model behave consistently.

@cr-composition

:::
